### 2022-10-04 PSTAT131_Final_Project.Rmd
progress

**Loading the dataset:**  
```{r}
pokemon <- read.csv("pokemon.csv")
head(pokemon)
```

Outcome-  
capture_rate: capture rate

https://bulbapedia.bulbagarden.net/wiki/Stat
Variables of interest-  
attack (how much damage they deal using a physical move)  
classification 
defense 
hp (hit points - how much damage they can receive before fainting)
sp_attack (how much damage the deal using special moves)   
sp_defense  
type1  
type2  
generation
is_legendary

```{r}
library(tidyverse)
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
```

```{r}

library(dplyr)
filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
pokemon[pokemon$name=="Minior", "capture_rate"] <- 30
# I chose 30 because the capture rate only rises to 255 if you break its shell
filter(pokemon, name %in% c("Minior"))
```

```{r}
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
pokemon[, "capture_rate"] <- as.numeric(pokemon$capture_rate)
```

```{r}
ggplot(pokemon, aes(x=attack, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=hp, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=generation, y=capture_rate, group=generation)) + geom_boxplot()
ggplot(pokemon, aes(x=defense, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=sp_attack, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=sp_defense, y=capture_rate)) + geom_point()
ggplot(pokemon, aes(x=classfication, y=capture_rate, group=classfication)) + geom_boxplot()
ggplot(pokemon, aes(x=type1, y=capture_rate, group=type2)) + geom_boxplot()
```

### 2022-10-04 PSTAT131_Final_Project.Rmd
edit

@@
 ```
+
+```{r}
+pokemon %>%
+  distinct(classfication)
+```
+
+Might not be a good idea to do classification because there are 588 unique classifications of a pokemon --> better to focus on type
+
+```{r}
+library(corrplot)
+pokemon_numeric <- pokemon %>%
+  select_if(is.numeric)
+# taking out the aforementioned columns with too many missing values (was not planning on using them anyways)
+pokemon_numeric <- pokemon_numeric[, !names(pokemon_numeric) %in% c("height_m", "weight_kg", "percentage_male")]
+
+pokemon_cor <- cor(pokemon_numeric)
+pokemon_corrplot <- corrplot(pokemon_cor)
+```
+
+We can see that how a pokemon performs against a certain other type of pokemon is not really correlated to capture rate much.
+However, this also shows how capture rate has a high negative correlation with base_total which is the Base Stat Total (BST = hp + attack + defense + sp_attack + sp_defense + speed).
+So, it might be worth exploring how all of those added together affects the capture rate.
+
+```{r}
+ggplot(pokemon, aes(capture_rate)) + geom_histogram()
+```
+
+No specific skewness is visible, but there is a peak in the data around 40-50 capture rate.

### 2022-10-08 PSTAT131_Final_Project.Rmd
edit

@@
+Loading, exploring, morphing  --> tidying eventually where we make factors
+
 **Loading the dataset:**  
 ```{r}
+library(tidyverse)
 pokemon <- read.csv("pokemon.csv")
@@
 ```
+
+```{r}
+pokemon %>%
+  dim()
+```
+801 rows - 801 pokemon
+41 columns - we definitely will not be needing all of those variables
+
+based on the question i want to answer, it would not be very significant to look at the attack scores of pokemon against other specific types of pokemon as we already have the overall attack score.
+also it would not be meaningful to do based on classification as there are N unique classifications (use code from below).
+base egg steps and base happiness appear to be about the same for most pokemon and do not really have much to do with the skill level of the pokemon.
+experience growth also appears to be quite similar across most pokemon so we are also not interested in that.
+
+further exploring the capture rate, our outcome. notice that one is a string --> minior --> explain why choose 30 instead of 255
 

### 2022-10-08 Draft.Rmd
edit

## Loading and Exploring Raw Data

## Morphing and Tidying Data

### 2022-10-08 Draft.Rmd
loading and exploring

@@
 
+```{r}
+# Loading all necessary packages
+library(tidyverse)
+library(dplyr)
+library(tidymodels)
+```
+
+Now, let's load our data and start our exploration to see what we're working with!
+
+```{r}
+# Assigning the data to a variable
+pokemon <- read.csv("data/pokemon.csv")
+
+# Calling head() to see the first few rows
+head(pokemon)
+```
+
+How many rows and columns are we working with?
+
+```{r}
+pokemon %>%
+  dim()
+```
+
+We have a total of 801 rows, one for each Pokemon in the dataset, and 41 columns, which are the attributes reported about each Pokemon.
+Within these attributes, we have our response variable, `capture_rate`, as well as 40 features.
+Obviously, we will not be working with all 40 features so we will need to continue exploring the variables to decide which ones should be kept and which should not.
+
+From the data, each of the variables that specify each Pokemon's skill against a certain type of Pokemon (i.e.
+`against_bug`, `against_dark`, etc.) do not appear to be as relevant as other attributes in deciding whether specifc features about a Pokemon affect its capture rate.
+We can also assume that the `attack` variable gives us an overall estimate of each Pokemon's attack skill so we would not need to go into such great detail as finding out how the Pokemon perform against every other type of Pokemon.
+For that reason, these 18 columns can be dropped.
+
+At first, it seems that the `classification` variable may be helpful; however, let us see how many unique classifications of a Pokemon there are.
+
+```{r}
+pokemon %>%
+  distinct(classfication) %>%
+  count()
+```
+
+There are 588 unique classifications! That seems a little too high for our analysis so we will just stick to using the `type` of the Pokemon instead.
+
+Furthermore, for the `abilities` variable, we can see that there is no set amount for how many abilities a Pokemon has or a measure to figure out which abilities are more powerful than others.
+For these reasons, it may be best to avoid including this column into our analysis.
+
+Also, the `base_happiness`, `base_egg_steps`, and `experience_growth` variables appear to be the same across most Pokemon as they are base statistics that each Pokemon shares.
+These variables would not be helpful in our analysis since we are trying to find how the major differences between each Pokemon affect their capture rate.
+
+Some other variables must also be taken out as they do not seem as if they would reveal any more information relevant to our analysis: `weight_kg`, `pokedex_number`, `percentage_male`, `japanese_name`, and `height_m`.
+
 ## Morphing and Tidying Data

### 2022-10-20 Draft2.Rmd
progress

```{r}
# Loading all necessary packages
library(tidyverse)
library(dplyr)
library(tidymodels)
library(readr)
# Assigning the data to a variable
pokemon <- read_csv("data/pokemon.csv")
```

```{r}
# Calling head() to see the first few rows
head(pokemon)
```

```{r}
# Calling dim() to see how many rows and columns
dim(pokemon)
```

```{r}
# Selecting only the variables we want
pokemon <- pokemon %>% 
  select(c("abilities", "attack", "base_total", "capture_rate", "defense", "hp", "name", "sp_attack", "sp_defense", "speed", "type1", "generation", "is_legendary"))
head(pokemon)
```

```{r}
# summary stats
pokemon %>% 
  summary()

# missing values?
colSums(is.na(pokemon))
```

```{r}
# capture rate should not be a character
library(dplyr)
filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
pokemon[pokemon$name=="Minior", "capture_rate"] <- "30"
pokemon$capture_rate <- as.integer(pokemon$capture_rate)
# I chose 30 because the capture rate only rises to 255 if you break its shell
filter(pokemon, name %in% c("Minior"))
```

```{r}
generation_levels <- c(1,2,3,4,5,6,7)
pokemon$generation <- factor(pokemon$generation, levels=generation_levels)
```

```{r}
pokemon %>% 
  summary()
```

### 2022-10-20 Editing_and_Exploring.Rmd
edit
progress

--- 2022-10-08 PSTAT131_Final_Project.Rmd
@@
 library(tidyverse)
-pokemon <- read.csv("pokemon.csv")
+pokemon <- read.csv("data/pokemon.csv")
 head(pokemon)
@@
 No specific skewness is visible, but there is a peak in the data around 40-50 capture rate.
+
+```{r}
+
+ability_split <-strsplit(as.character(pokemon$abilities),',',fixed=TRUE)
+#ability_split
+```
+
+```{r}
+ability_df <- tibble(line = 1:801, text = pokemon$abilities)
+ability_df
+```

### 2022-10-23 Draft2.Rmd
progress

@@
 pokemon$generation <- factor(pokemon$generation, levels=generation_levels)
+# MAKE IS_LEGENDARY A FACTOR
 ```
@@
 ```
+
+```{r}
+library(tidytext)
+```
+
+```{r}
+test <- pokemon$abilities %>% 
+  str_split(pattern = "', '") #%>% 
+  # str_extract_all(string = ., pattern = "\\w+", simplify = TRUE)
+  # map(str_extract_all(string = ., pattern = "\\w+", simplify = TRUE))
+all_abilities <- test %>% map(~ str_replace_all(string = ., pattern = "[[:punct:]]", "")) %>% 
+  unlist() %>% as.factor() %>% 
+  data.frame(abilities = .)
+
+all_abilities %>% group_by(abilities) %>% summarise(count = n()) %>% 
+  arrange(desc(count)) %>% View()
+
+ggplot(data = tibble(all_abilities)) +
+  geom_bar(aes(x = all_abilities))
+```
+
+```{r}
+pokemon_copy <- pokemon
+pokemon_copy$abilities <- pokemon_copy$abilities %>% 
+  str_split(pattern = "', '") %>% 
+  map(~ str_replace_all(string = ., pattern = "[[:punct:]]", ""))
+
+#tidy_abilities <- pokemon_copy %>% 
+#  unnest_tokens(words, abilities)
+
+pokemon_copy %>% 
+  unnest() %>% 
+  View()
+```

### 2022-10-23 Draft2.Rmd
progress

@@
 
-ggplot(data = tibble(all_abilities)) +
-  geom_bar(aes(x = all_abilities))
+#ggplot(data = tibble(all_abilities)) +
+ # geom_bar(aes(x = all_abilities))
 ```
@@
 
-pokemon_copy %>% 
-  unnest() %>% 
-  View()
+pokemon_copy2 <- pokemon_copy %>% 
+  unnest()
+
+pokemon_copy2
 ```
+
+```{r}
+library(dplyr)
+
+pokemon_abilities_grouped <- pokemon_copy2 %>% 
+  group_by(abilities)
+
+pokemon_abilities_grouped %>% 
+  summarise(
+    capture_rate = mean(capture_rate)
+  ) %>% 
+  arrange(desc(capture_rate))
+
+```

### 2022-10-26 Draft2.Rmd
progress

@@
 
-all_abilities %>% group_by(abilities) %>% summarise(count = n()) %>% 
+all_abilities %>% group_by(abilities) %>% summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
   arrange(desc(count)) %>% View()
@@
 
-pokemon_copy2
+pokemon_copy2 %>% group_by(abilities) %>%  summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
+  arrange(desc(count)) %>% View()
 ```
@@
 ```
+
+```{r}
+mean(pokemon_copy2$capture_rate)
+```

### 2022-11-07 Draft2.Rmd
edit

@@
 pokemon <- read_csv("data/pokemon.csv")
+tidymodels_prefer()
 ```
@@
 ```
+
+```{r}
+o
+```

### 2022-11-07 Draft2.Rmd
progress

--- 2022-10-26 Draft2.Rmd
@@
 pokemon <- read_csv("data/pokemon.csv")
+tidymodels_prefer()
+set.seed(0124)
 ```
@@
 ```
+
+```{r}
+# making boolean columns for if a pokemon has the top 5 most apparent abilities
+pokemon_copy3 <- pokemon
+
+sturdy <- "Sturdy"
+pokemon_copy4 <- pokemon %>% 
+  mutate(has_sturdy = if_else(grepl("Sturdy", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_swift_swim = if_else(grepl("Swift Swim", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_keen_eye = if_else(grepl("Keen Eye", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_chlorophyll = if_else(grepl("Chlorophyll", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_levitate = if_else(grepl("Levitate", pokemon$abilities), 1, 0))
+
+sum(pokemon_copy4$has_sturdy)
+sum(pokemon_copy4$has_swift_swim)
+sum(pokemon_copy4$has_keen_eye)
+sum(pokemon_copy4$has_chlorophyll)
+sum(pokemon_copy4$has_levitate)
+
+write_csv(pokemon_copy4, "pokemon_final.csv")
+```
+
+```{r}
+pokemon_data <- read_csv("pokemon_final.csv")
+head(pokemon_data)
+```
+
+```{r}
+# Changing categorical predictors to factors
+pokemon_data$type1 <- as.factor(pokemon_data$type1)
+pokemon_data$generation <- as.factor(pokemon_data$generation)
+pokemon_data$is_legendary <- as.factor(pokemon_data$is_legendary)
+pokemon_data$has_sturdy <- as.factor(pokemon_data$has_sturdy)
+pokemon_data$has_swift_swim <- as.factor(pokemon_data$has_swift_swim)
+pokemon_data$has_keen_eye <- as.factor(pokemon_data$has_keen_eye)
+pokemon_data$has_chlorophyll <- as.factor(pokemon_data$has_chlorophyll)
+pokemon_data$has_levitate <- as.factor(pokemon_data$has_levitate)
+```
+
+```{r}
+summary(pokemon_data)
+```
+
+```{r}
+# Splitting the data
+pokemon_split <- initial_split(pokemon_data, prop = 0.7, strata = capture_rate)
+pokemon_train <- training(pokemon_split)
+pokemon_test <- testing(pokemon_split)
+```
+
+```{r}
+# Creating recipe
+pokemon_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+  step_dummy(all_nominal_predictors()) %>% 
+  step_center(all_predictors()) %>% 
+  step_scale(all_predictors())
+```
+
+```{r}
+# Creating folds
+pokemon_folds <- vfold_cv(pokemon_train, v = 10)
+```
+
+```{r}
+# Linear regression 
+lm_model <- linear_reg() %>% 
+  set_engine("lm")
+
+lm_workflow <- workflow() %>% 
+  add_model(lm_model) %>% 
+  add_recipe(pokemon_recipe)
+```
+
+```{r}
+# Fitting the models
+lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
+```
+```{r}
+collect_metrics(lm_fit)
+```

### 2022-11-14 Draft2.Rmd
progress

@@
 library(readr)
+library(kknn)
 # Assigning the data to a variable
@@
 
+### Linear Regression
+
 ```{r}
@@
 ```
+
+### Polynomial Regression
+
+### Ridge Regression   
+```{r}
+ridge_lasso_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+  step_dummy(all_nominal_predictors()) %>% 
+  step_novel(all_nominal_predictors()) %>% 
+  step_zv(all_predictors()) %>% 
+  step_normalize(all_predictors())
+
+ridge_spec <- linear_reg(mixture = 0, penalty = tune()) %>% 
+  set_mode("regression") %>% 
+  set_engine("glmnet")
+
+ridge_workflow <- workflow() %>% 
+  add_recipe(ridge_lasso_recipe) %>% 
+  add_model(ridge_spec)
+
+penalty_grid <- grid_regular(penalty(range = c(-5,5)), levels = 50)
+
+tune_res <- tune_grid(
+  ridge_workflow,
+  resamples = pokemon_folds,
+  grid = penalty_grid
+)
+```
+```{r}
+autoplot(tune_res)
+```
+
+```{r}
+best_penalty <- select_best(tune_res, metric = "rsq")
+
+ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
+ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
+collect_metrics(ridge_final_fit)
+```
+
+### Lasso Regression 
+```{r}
+lasso_spec <- 
+  linear_reg(penalty = tune(), mixture = 1) %>% 
+  set_mode("regression") %>% 
+  set_engine("glmnet")
+
+lasso_workflow <- workflow() %>% 
+  add_recipe(ridge_lasso_recipe) %>% 
+  add_model(lasso_spec)
+
+tune_res2 <- tune_grid(
+  lasso_workflow,
+  resamples = pokemon_folds,
+  grid = penalty_grid
+)
+```
+
+```{r}
+autoplot(tune_res2)
+```
+
+```{r}
+best_lasso_penalty <- select_best(tune_res2, metric = "rsq")
+lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
+lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
+```
+
+```{r}
+collect_metrics(lasso_final_fit)
+```
+
+```{r, message = FALSE, warning = FALSE}
+poly_rec <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+  step_dummy(all_nominal_predictors()) %>%
+  step_poly(all_numeric_predictors(), degree = tune())
+
+#step_poly(sp_attack, degree = 2) %>% 
+ # step_poly(defense, degree = 2) %>% 
+  #step_poly(sp_defense, degree = 2) %>% 
+#  step_poly(hp, degree = 2) %>% 
+ # step_poly(speed, degree = 2)
+
+lm_spec <- linear_reg() %>% 
+  set_mode("regression") %>% 
+  set_engine("lm")
+
+poly_wf <- workflow() %>% 
+  add_model(lm_spec) %>% 
+  add_recipe(poly_rec)
+
+degree_grid <- grid_regular(degree(range = c(1,5)), levels = 5)
+
+tune_res3 <- tune_grid(
+  poly_wf,
+  resamples = pokemon_folds,
+  grid = degree_grid
+)
+autoplot(tune_res3)
+```
+
+### KNN  
+```{r}
+knn_model <- nearest_neighbor(
+  neighbors = tune(),
+  mode = "regression") %>% 
+  set_engine("kknn")
+
+knn_workflow <- workflow() %>% 
+  add_model(knn_model) %>% 
+  add_recipe(pokemon_recipe)
+
+knn_parameters <- parameters(knn_model)
+
+knn_grd <- grid_regular(knn_parameters, levels = 5)
+
+knn_tune <- tune_grid(
+    knn_workflow,
+    resamples = pokemon_folds,
+    grid = knn_grd
+  )
+```
+
+```{r}
+best_knn <- select_best(knn_tune, metric = "rmse")
+knn_final <- finalize_workflow(knn_workflow, best_knn)
+knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
+```
+
+```{r}
+collect_metrics(knn_final_fit)
+```

### 2022-11-14 Final_Draft.Rmd
progress

---
title: "Predicting the Capture Rate of Pokemon"
subtitle: "PSTAT 131 Final Project"
author: "Shivani Kharva"
date: "2022-12-06"
output:
  html_document:
    code_folding: hide
    toc: true
---

## Introduction

The purpose of this project is to develop a model that will predict the capture rate of a Pokemon.

### What Are Pokemon?

Pokemon (or "pocket monsters") are creatures from the Japanese Nintendo video game, Pokemon. Pokemon is a role-playing game in which players play as trainers who capture the Pokemon using Poke Balls and train them to compete in battles with other trainers. As they train and gain more experience through battle, Pokemon evolve and grow stronger. There are over 800 Pokemon with 18 types (i.e. fire, water, etc,), and they have numerous abilities (both common and unique).

### What Are We Trying to Do?

However, a question I'd like to pose is "How hard is to capture a Pokemon in the first place depending on their existing statistics?" It seems quite reasonable to hypothesize that a Pokemon's capture rate would vary depending on variables like their type, base statistics, generation, etc. So, I would like to test if it is possible to develop a model that can predict the capture rate of a Pokemon based on those qualities.

### Why?

Players, including me, are often only able to see the aforementioned qualities of a Pokemon and not their actual capture rate. So, it may be useful to have a specific model that can accurately predict the capture rate of a Pokemon so a player may see how difficult it may be to capture that Pokemon. The player can then adjust their strategy if a Pokemon is harder to catch by using more advanced equipment, such as using an Ultra Ball instead of a Poke Ball or giving the Pokemon a fruit before attempting to capture them to slightly lower their guard. So, let's see if the variation in the capture rate of a specific Pokemon can be explained by these qualities!

## Loading Packages and Data

Let's see what data we're currently working with!

```{r, message = FALSE}
# Loading all necessary packages
library(tidyverse)
library(dplyr)
library(tidymodels)
library(readr)
library(kknn)

# Assigning the data to a variable
pokemon <- read_csv("data/pokemon.csv")
tidymodels_prefer()

# Setting the seed
set.seed(0124)
```

## Loading and Exploring the Raw Data

```{r}
# Calling head() to see the first few rows
head(pokemon)
```

This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.

```{r}
# Calling dim() to see how many rows and columns
dim(pokemon)
```

The data set contains 801 rows and 41 columns. This means we have 801 Pokemon and 41 variables. One of those variables, `capture_rate`, is our response, but that leaves us with 40 predictor variables! We're definitely going to have to narrow that down!

### Sources

### 2022-11-15 Final_Draft.Rmd
progress

@@
 library(kknn)
+library(dplyr)
 
@@
 
-## Loading and Exploring the Raw Data
+## Loading, Exploring, and Cleaning the Raw Data
 
@@
 
+From the data, each of the variables that specify each Pokemon's skill against a certain type of Pokemon (i.e.
+`against_bug`, `against_dark`, etc.) do not appear to be as relevant as other attributes in deciding whether specifc features about a Pokemon affect its capture rate.
+We can also assume that the `attack` variable gives us an overall estimate of each Pokemon's attack skill so we would not need to go into such great detail as finding out how the Pokemon perform against every other type of Pokemon.
+For that reason, these 18 columns can be dropped.
+
+At first, it seems that the `classification` variable may be helpful; however, let us see how many unique classifications of a Pokemon there are.
+
+```{r}
+pokemon %>%
+  distinct(classfication) %>%
+  count()
+```
+
+There are 588 unique classifications! That seems a little too high for our analysis so we will just stick to using the `type` of the Pokemon instead.
+
+Furthermore, for the `abilities` variable, we can see that there is no set amount for how many abilities a Pokemon has or a measure to figure out which abilities are more powerful than others.
+For these reasons, it may be best to avoid including this column into our analysis.
+
+Some other variables must also be taken out as they do not seem as if they would reveal any more information relevant to our analysis based on substantive knowledge: `weight_kg`, `pokedex_number`, `percentage_male`, `japanese_name`, `base_happiness`, `base_egg_steps`, `experience_growth`, and `height_m`.
+
+Now, let's filter out all those unwanted variables!!
+
+```{r}
+# Selecting only the variables we want
+pokemon <- pokemon %>% 
+  select(c("abilities", "attack", "base_total", "capture_rate", "defense", "hp", "name", "sp_attack", "sp_defense", "speed", "type1", "generation", "is_legendary"))
+head(pokemon)
+```
+
 ### Sources

### 2022-11-22 Draft2.Rmd
progress

@@
 library(kknn)
+library(janitor)
 # Assigning the data to a variable
@@
 pokemon <- pokemon %>% 
-  select(c("abilities", "attack", "base_total", "capture_rate", "defense", "hp", "name", "sp_attack", "sp_defense", "speed", "type1", "generation", "is_legendary"))
+  select(c("abilities", "attack", "base_happiness",  "base_total", "capture_rate", "defense", "experience_growth", "height_m", "percentage_male", "hp", "name", "sp_attack", "sp_defense", "speed", "type1", "weight_kg", "generation", "is_legendary")) %>%
+  clean_names()
 head(pokemon)
@@
 # Creating recipe
-pokemon_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
-  step_dummy(all_nominal_predictors()) %>% 
+pokemon_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+  # imputing percentage male
+  step_impute_linear(percentage_male, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  # imputing height
+  step_impute_linear(height_m, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  # imputing weight
+  step_impute_linear(weight_kg, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, height_m, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_other(type1) %>% 
+  # dummy coding nominal variables
+  step_dummy(generation, is_legendary, type1, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate) %>% 
+  # normalizing
   step_center(all_predictors()) %>% 
@@
 # Creating folds
-pokemon_folds <- vfold_cv(pokemon_train, v = 10)
+pokemon_folds <- vfold_cv(pokemon_train, v = 10, strata = capture_rate)
 ```
@@
 ```
+
 ```{r}
@@
 ```{r}
-ridge_lasso_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
-  step_dummy(all_nominal_predictors()) %>% 
+ridge_lasso_recipe <- pokemon_recipe %>% 
   step_novel(all_nominal_predictors()) %>% 
@@
 ```{r}
-best_lasso_penalty <- select_best(tune_res2, metric = "rsq")
+best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
 lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
@@
 ```{r, message = FALSE, warning = FALSE}
-poly_rec <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
-  step_dummy(all_nominal_predictors()) %>%
-  step_poly(all_numeric_predictors(), degree = tune())
+poly_rec <- pokemon_recipe %>% 
+  step_poly(attack, defense, hp, sp_attack, sp_defense, speed, degree = tune())
+
+ggplot(pokemon_train, aes(y = type1)) + geom_bar()
 
@@
 
+```{r}
+best_poly<- select_best(tune_res3, metric = "rmse")
+poly_final <- finalize_workflow(poly_wf, best_poly)
+poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
+
+collect_metrics(poly_final_fit)
+```
+
 ### KNN  
@@
 ```{r}
+autoplot(knn_tune)
+```
+
+```{r}
 best_knn <- select_best(knn_tune, metric = "rmse")
@@
 ```
+
+### Elastic Net
+
+```{r, warning = FALSE, message = FALSE}
+elastic_spec <- 
+  linear_reg(penalty = tune(), mixture = tune()) %>% 
+  set_mode("regression") %>% 
+  set_engine("glmnet")
+
+elastic_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(elastic_spec)
+
+elastic_grid <- grid_regular(penalty(range = c(-5, 5)), mixture(range = c(0,1)), levels = 10)
+
+elastic_tune <- tune_grid(
+  elastic_workflow,
+  resamples = pokemon_folds,
+  grid = elastic_grid
+)
+```
+
+```{r}
+autoplot(elastic_tune)
+```
+
+```{r}
+best_elastic <- select_best(elastic_tune, metric = "rmse")
+elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
+elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
+
+collect_metrics(elastic_final_fit)
+```

### 2022-11-25 Draft2.Rmd
sad times

@@
 library(janitor)
+library(ISLR)
+library(discrim)
+library(poissonreg)
+library(glmnet)
+library(corrr)
+library(corrplot)
+library(randomForest)
+library(xgboost)
+library(rpart.plot)
+library(vip)
+library(ranger)
 # Assigning the data to a variable
@@
 ```
+
+```{r}
+# Setting up random forest model with specified tuning parameters and engine
+rf_spec <- rand_forest(mtry = tune(), trees = tune(), min_n = tune()) %>% 
+  set_engine("ranger", importance = "impurity") %>% 
+  set_mode("regression")
+
+# Setting up random forest workflow
+rf_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(rf_spec)
+
+# Creating a regular grid with 8 levels each
+rf_parameter_grid <- grid_regular(mtry(range = c(1, 19)), trees(range = c(200,1000)), min_n(range = c(5,20)), levels = 8)
+```
+
+```{r, eval = FALSE}
+# Tuning the model with roc_auc as a metric
+rf_tune_res <- tune_grid(
+  rf_workflow,
+  resamples = pokemon_folds,
+  grid = rf_parameter_grid
+)
+
+# Saving the fit
+write_rds(rf_tune_res, file = "rf.rds")
+```
+
+```{r}
+# Loading in the random forest
+rf_tuned <- read_rds(file = "rf.rds")
+
+# Producing autoplot of results
+autoplot(rf_tuned)
+```
+
+```{r}
+# Finding roc_auc of best performing random forest model on folds
+rf_best_roc_auc <- rf_tuned %>% 
+  collect_metrics() %>% 
+  arrange(desc(mean)) %>% 
+  head(1)
+
+rf_best_roc_auc
+```
+
+```{r}
+best_rf <- select_best(rf_tuned)
+rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
+rf_final_fit <- fit(rf_final_workflow, data = pokemon_train)
+
+rf_final_fit %>% 
+  extract_fit_engine() %>% 
+  vip()
+```
+
+### BOOSTED TREE
+```{r}
+# Setting up boosted tree model
+boosted_spec <- boost_tree(trees = tune()) %>%
+  set_engine("xgboost") %>%
+  set_mode("regression")
+
+# Setting up boosted trees workflow
+boosted_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(boosted_spec)
+
+# Creating regular grid
+boosted_grid <- grid_regular(trees(range = c(10, 2000)), levels = 10)
+```
+
+```{r, eval = FALSE}
+# Tuning and specifying roc_auc
+boosted_tune_res <- tune_grid(
+  boosted_workflow,
+  resamples = pokemon_folds,
+  grid = boosted_grid
+)
+
+# Saving the fit
+write_rds(boosted_tune_res, file = "boosted.rds")
+```
+
+```{r}
+# Loading the fit
+boosted_tuned <- read_rds(file = "boosted.rds")
+
+# Printing autoplot of results
+autoplot(boosted_tuned)
+```
+
+```{r}
+# Finding roc_auc of best performing boosted tree model on folds
+boosted_best_roc_auc <- boosted_tuned %>% 
+  collect_metrics() %>% 
+  arrange(desc(mean)) %>% 
+  head(1)
+
+boosted_best_roc_auc
+```

### 2022-11-30 Draft2.Rmd
commit

@@
 ```{r}
+ggplot(pokemon_data, aes(capture_rate)) + 
+  geom_bar(aes(fill = is_legendary))
+
+ggplot(pokemon_data, aes(capture_rate)) + 
+  geom_bar(aes(fill = generation))
+```
+
+```{r}
 summary(pokemon_data)
@@
 
-ggplot(pokemon_train, aes(y = type1)) + geom_bar()
+#ggplot(pokemon_train, aes(y = type1)) + geom_bar()
 
@@
   collect_metrics() %>% 
-  arrange(desc(mean)) %>% 
-  head(1)
+  arrange(mean)
 
@@
   collect_metrics() %>% 
-  arrange(desc(mean)) %>% 
-  head(1)
+  arrange((mean))
 

### 2022-11-30 Final_Draft.Rmd
commit

@@
 
-Let's see what data we're currently working with!
-
-```{r, message = FALSE}
+```{r, message = FALSE, warning = FALSE}
 # Loading all necessary packages
@@
 library(kknn)
-library(dplyr)
+library(janitor)
+library(ISLR)
+library(discrim)
+library(poissonreg)
+library(glmnet)
+library(corrr)
+library(corrplot)
+library(randomForest)
+library(xgboost)
+library(rpart.plot)
+library(vip)
+library(ranger)
 
@@
 tidymodels_prefer()
+set.seed(0124)
 
-# Setting the seed
-set.seed(0124)
-```
-
-## Loading, Exploring, and Cleaning the Raw Data
-
-```{r}
 # Calling head() to see the first few rows
@@
 This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.
+
+## Exploring and Cleaning the Raw Data
+
+Let's mess around with the data a little bit to see what we're currently working with!
 
@@
 
-Some other variables must also be taken out as they do not seem as if they would reveal any more information relevant to our analysis based on substantive knowledge: `weight_kg`, `pokedex_number`, `percentage_male`, `japanese_name`, `base_happiness`, `base_egg_steps`, `experience_growth`, and `height_m`.
+The `base_egg_steps` variable is a variable not used in the original Pokemon games so, to keep the data consistent with the game it was taken from, this variable should also be omitted.
+
+Finally, `pokedex_number` and `japanese_name` are unique to each Pokemon and are mainly used to identify the Pokemon; however, since we already have the `name` variable, it would not be useful to include these variables.
 
@@
 
+Upon further inspection, we can see that our outcome, `capture_rate`, is actually a character variable, which is not what we want.
+By looking into the values, we see that one of the Pokemon, Minior, actually has two capture rates.
+
+```{r}
+# Finding which Pokemon has a character capture rate
+filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
+```
+
+By looking at [Bulbapedia](https://bulbapedia.bulbagarden.net/wiki/Minior_(Pok%C3%A9mon)), a community Pokemon encyclopedia, we can see that Minior has a relatively low capture rate at 30 when it is "Meteorite," meaning its shell has not been broken.
+However, once Minior's shell is broken ("Core"), the Pokemon shoots up to a capture rate of 255, which is the highest capture rate there is.
+Therefore, it would make sense to keep the capture rate of Minior when it is still Meteorite.
+
+```{r}
+# Changing Minior to capture rate 30
+pokemon[pokemon$name=="Minior", "capture_rate"] <- "30"
+
+# Making capture_rate an integer
+pokemon$capture_rate <- as.integer(pokemon$capture_rate)
+```
+
+Now our outcome is an integer we can use in regression!
+
 ### Sources

### 2022-11-30 testingsomething.Rmd
commit

--- 2022-11-22 Draft2.Rmd
@@
 set.seed(0124)
-```
-
-```{r}
-# Calling head() to see the first few rows
-head(pokemon)
-```
-
-```{r}
-# Calling dim() to see how many rows and columns
-dim(pokemon)
-```
-
-```{r}
-# Selecting only the variables we want
-pokemon <- pokemon %>% 
-  select(c("abilities", "attack", "base_happiness",  "base_total", "capture_rate", "defense", "experience_growth", "height_m", "percentage_male", "hp", "name", "sp_attack", "sp_defense", "speed", "type1", "weight_kg", "generation", "is_legendary")) %>%
-  clean_names()
-head(pokemon)
-```
-
-```{r}
-# summary stats
-pokemon %>% 
-  summary()
-
-# missing values?
-colSums(is.na(pokemon))
-```
-
-```{r}
-# capture rate should not be a character
-library(dplyr)
-filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
-pokemon[pokemon$name=="Minior", "capture_rate"] <- "30"
-pokemon$capture_rate <- as.integer(pokemon$capture_rate)
-# I chose 30 because the capture rate only rises to 255 if you break its shell
-filter(pokemon, name %in% c("Minior"))
-```
-
-```{r}
-generation_levels <- c(1,2,3,4,5,6,7)
-pokemon$generation <- factor(pokemon$generation, levels=generation_levels)
-# MAKE IS_LEGENDARY A FACTOR
-```
-
-```{r}
-pokemon %>% 
-  summary()
-```
-
-```{r}
-library(tidytext)
-```
-
-```{r}
-test <- pokemon$abilities %>% 
-  str_split(pattern = "', '") #%>% 
-  # str_extract_all(string = ., pattern = "\\w+", simplify = TRUE)
-  # map(str_extract_all(string = ., pattern = "\\w+", simplify = TRUE))
-all_abilities <- test %>% map(~ str_replace_all(string = ., pattern = "[[:punct:]]", "")) %>% 
-  unlist() %>% as.factor() %>% 
-  data.frame(abilities = .)
-
-all_abilities %>% group_by(abilities) %>% summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
-  arrange(desc(count)) %>% View()
-
-#ggplot(data = tibble(all_abilities)) +
- # geom_bar(aes(x = all_abilities))
-```
-
-```{r}
-pokemon_copy <- pokemon
-pokemon_copy$abilities <- pokemon_copy$abilities %>% 
-  str_split(pattern = "', '") %>% 
-  map(~ str_replace_all(string = ., pattern = "[[:punct:]]", ""))
-
-#tidy_abilities <- pokemon_copy %>% 
-#  unnest_tokens(words, abilities)
-
-pokemon_copy2 <- pokemon_copy %>% 
-  unnest()
-
-pokemon_copy2 %>% group_by(abilities) %>%  summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
-  arrange(desc(count)) %>% View()
-```
-
-```{r}
-library(dplyr)
-
-pokemon_abilities_grouped <- pokemon_copy2 %>% 
-  group_by(abilities)
-
-pokemon_abilities_grouped %>% 
-  summarise(
-    capture_rate = mean(capture_rate)
-  ) %>% 
-  arrange(desc(capture_rate))
-
-```
-
-```{r}
-mean(pokemon_copy2$capture_rate)
-```
-
-```{r}
-# making boolean columns for if a pokemon has the top 5 most apparent abilities
-pokemon_copy3 <- pokemon
-
-sturdy <- "Sturdy"
-pokemon_copy4 <- pokemon %>% 
-  mutate(has_sturdy = if_else(grepl("Sturdy", pokemon$abilities), 1, 0)) %>% 
-  mutate(has_swift_swim = if_else(grepl("Swift Swim", pokemon$abilities), 1, 0)) %>% 
-  mutate(has_keen_eye = if_else(grepl("Keen Eye", pokemon$abilities), 1, 0)) %>% 
-  mutate(has_chlorophyll = if_else(grepl("Chlorophyll", pokemon$abilities), 1, 0)) %>% 
-  mutate(has_levitate = if_else(grepl("Levitate", pokemon$abilities), 1, 0))
-
-sum(pokemon_copy4$has_sturdy)
-sum(pokemon_copy4$has_swift_swim)
-sum(pokemon_copy4$has_keen_eye)
-sum(pokemon_copy4$has_chlorophyll)
-sum(pokemon_copy4$has_levitate)
-
-write_csv(pokemon_copy4, "pokemon_final.csv")
 ```
@@
 ```{r}
-summary(pokemon_data)
-```
-
-```{r}
 # Splitting the data
@@
 # Creating recipe
-pokemon_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+pokemon_recipe <- recipe(capture_rate ~ base_total + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
   # imputing percentage male
-  step_impute_linear(percentage_male, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_impute_linear(percentage_male, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
   # imputing height
-  step_impute_linear(height_m, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_impute_linear(height_m, impute_with = imp_vars(base_total, base_happiness, experience_growth, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
   # imputing weight
-  step_impute_linear(weight_kg, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, height_m, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_impute_linear(weight_kg, impute_with = imp_vars(base_total, base_happiness, experience_growth, height_m, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
   step_other(type1) %>% 
@@
 ```
-
-### Linear Regression
 
@@
 
-### Polynomial Regression
-
-### Ridge Regression   
-```{r}
-ridge_lasso_recipe <- pokemon_recipe %>% 
-  step_novel(all_nominal_predictors()) %>% 
-  step_zv(all_predictors()) %>% 
-  step_normalize(all_predictors())
-
-ridge_spec <- linear_reg(mixture = 0, penalty = tune()) %>% 
-  set_mode("regression") %>% 
-  set_engine("glmnet")
-
-ridge_workflow <- workflow() %>% 
-  add_recipe(ridge_lasso_recipe) %>% 
-  add_model(ridge_spec)
-
-penalty_grid <- grid_regular(penalty(range = c(-5,5)), levels = 50)
-
-tune_res <- tune_grid(
-  ridge_workflow,
-  resamples = pokemon_folds,
-  grid = penalty_grid
-)
-```
-```{r}
-autoplot(tune_res)
-```
-
-```{r}
-best_penalty <- select_best(tune_res, metric = "rsq")
-
-ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
-ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
-collect_metrics(ridge_final_fit)
-```
-
-### Lasso Regression 
-```{r}
-lasso_spec <- 
-  linear_reg(penalty = tune(), mixture = 1) %>% 
-  set_mode("regression") %>% 
-  set_engine("glmnet")
-
-lasso_workflow <- workflow() %>% 
-  add_recipe(ridge_lasso_recipe) %>% 
-  add_model(lasso_spec)
-
-tune_res2 <- tune_grid(
-  lasso_workflow,
-  resamples = pokemon_folds,
-  grid = penalty_grid
-)
-```
-
-```{r}
-autoplot(tune_res2)
-```
-
-```{r}
-best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
-lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
-lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
-```
-
-```{r}
-collect_metrics(lasso_final_fit)
-```
-
 ```{r, message = FALSE, warning = FALSE}
 poly_rec <- pokemon_recipe %>% 
-  step_poly(attack, defense, hp, sp_attack, sp_defense, speed, degree = tune())
+  step_poly(base_total, base_happiness, experience_growth, height_m, percentage_male, weight_kg, degree = tune())
 
-ggplot(pokemon_train, aes(y = type1)) + geom_bar()
+#ggplot(pokemon_train, aes(y = type1)) + geom_bar()
 
@@
 
-degree_grid <- grid_regular(degree(range = c(1,5)), levels = 5)
+degree_grid <- grid_regular(degree(range = c(1,6)), levels = 6)
 
@@
 
-### KNN  
 ```{r}
-knn_model <- nearest_neighbor(
-  neighbors = tune(),
-  mode = "regression") %>% 
-  set_engine("kknn")
+# Setting up boosted tree model
+boosted_spec <- boost_tree(trees = tune()) %>%
+  set_engine("xgboost") %>%
+  set_mode("regression")
 
-knn_workflow <- workflow() %>% 
-  add_model(knn_model) %>% 
-  add_recipe(pokemon_recipe)
+# Setting up boosted trees workflow
+boosted_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(boosted_spec)
 
-knn_parameters <- parameters(knn_model)
+# Creating regular grid
+boosted_grid <- grid_regular(trees(range = c(10, 2000)), levels = 10)
+```
 
-knn_grd <- grid_regular(knn_parameters, levels = 5)
+```{r, eval = FALSE}
+# Tuning and specifying roc_auc
+boosted_tune_res <- tune_grid(
+  boosted_workflow,
+  resamples = pokemon_folds,
+  grid = boosted_grid
+)
 
-knn_tune <- tune_grid(
-    knn_workflow,
-    resamples = pokemon_folds,
-    grid = knn_grd
-  )
+# Saving the fit
+write_rds(boosted_tune_res, file = "boosted_2.rds")
 ```
@@
 ```{r}
-autoplot(knn_tune)
+# Loading the fit
+boosted_tuned <- read_rds(file = "boosted.rds")
+
+# Printing autoplot of results
+autoplot(boosted_tuned)
 ```
@@
 ```{r}
-best_knn <- select_best(knn_tune, metric = "rmse")
-knn_final <- finalize_workflow(knn_workflow, best_knn)
-knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
+# Finding roc_auc of best performing boosted tree model on folds
+boosted_best_roc_auc <- boosted_tuned %>% 
+  collect_metrics() %>% 
+  arrange(desc(mean)) %>% 
+  head(1)
+
+boosted_best_roc_auc
 ```
-
-```{r}
-collect_metrics(knn_final_fit)
-```
-
-### Elastic Net
-
-```{r, warning = FALSE, message = FALSE}
-elastic_spec <- 
-  linear_reg(penalty = tune(), mixture = tune()) %>% 
-  set_mode("regression") %>% 
-  set_engine("glmnet")
-
-elastic_workflow <- workflow() %>% 
-  add_recipe(pokemon_recipe) %>% 
-  add_model(elastic_spec)
-
-elastic_grid <- grid_regular(penalty(range = c(-5, 5)), mixture(range = c(0,1)), levels = 10)
-
-elastic_tune <- tune_grid(
-  elastic_workflow,
-  resamples = pokemon_folds,
-  grid = elastic_grid
-)
-```
-
-```{r}
-autoplot(elastic_tune)
-```
-
-```{r}
-best_elastic <- select_best(elastic_tune, metric = "rmse")
-elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
-elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
-
-collect_metrics(elastic_final_fit)
-```

### 2022-12-07 Draft2.Rmd
commit

@@
 # Creating recipe
-pokemon_recipe <- recipe(capture_rate ~ attack + defense + hp + sp_attack + sp_defense + speed + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+pokemon_recipe <- recipe(capture_rate ~ base_total + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
   # imputing percentage male
-  step_impute_linear(percentage_male, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_impute_linear(percentage_male, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
   # imputing height
-  step_impute_linear(height_m, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_impute_linear(height_m, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
   # imputing weight
-  step_impute_linear(weight_kg, impute_with = imp_vars(attack, defense, sp_attack, sp_defense, speed, base_happiness, experience_growth, height_m, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  step_impute_linear(weight_kg, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
   step_other(type1) %>% 
@@
 ```{r}
-best_penalty <- select_best(tune_res, metric = "rsq")
+best_penalty <- select_best(tune_res, metric = 'rmse')
 
@@
 # Producing autoplot of results
-autoplot(rf_tuned)
+autoplot(rf_tuned, metric = 'rmse')
 ```

### 2022-12-07 testingsomething.Rmd
commit

@@
 pokemon_test <- testing(pokemon_split)
+
+#sd(pokemon_train$capture_rate)
 ```
@@
 # Creating recipe
-pokemon_recipe <- recipe(capture_rate ~ base_total + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
-  # imputing percentage male
-  step_impute_linear(percentage_male, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
-  # imputing height
-  step_impute_linear(height_m, impute_with = imp_vars(base_total, base_happiness, experience_growth, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
-  # imputing weight
-  step_impute_linear(weight_kg, impute_with = imp_vars(base_total, base_happiness, experience_growth, height_m, percentage_male, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
-  step_other(type1) %>% 
+pokemon_recipe <- recipe(capture_rate ~ base_total + is_legendary, data = pokemon_train) %>% 
   # dummy coding nominal variables
-  step_dummy(generation, is_legendary, type1, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate) %>% 
+  step_dummy(is_legendary) %>% 
   # normalizing
@@
 # Setting up boosted tree model
-boosted_spec <- boost_tree(trees = tune()) %>%
+boosted_spec <- boost_tree(trees = tune(),
+                          learn_rate = tune(),
+                          min_n = tune()) %>%
   set_engine("xgboost") %>%
@@
 # Creating regular grid
-boosted_grid <- grid_regular(trees(range = c(10, 2000)), levels = 10)
+boosted_grid <- grid_regular(trees(range = c(5, 200)), learn_rate(range = c(0.1,1), trans = identity_trans()), min_n(range = c(40, 60)), levels = 2)
 ```
@@
 # Saving the fit
-write_rds(boosted_tune_res, file = "boosted_2.rds")
+write_rds(boosted_tune_res, file = "boosted_3.rds")
 ```
@@
 # Loading the fit
-boosted_tuned <- read_rds(file = "boosted.rds")
+boosted_tuned <- read_rds(file = "boosted_3.rds")
 

### 2022-12-10 Draft2.Rmd
commit

@@
   step_impute_linear(weight_kg, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  # Only prioritizing top types of Pokemon
   step_other(type1) %>% 
@@
 lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
+
+# Save
+write_rds(lm_fit, file = "lm_final.rds")
 ```
@@
 ```{r}
-collect_metrics(lm_fit)
+lm_final_fit <- read_rds(file = "lm_final.rds")
+
+collect_metrics(lm_final_fit)
 ```
@@
 ridge_workflow <- workflow() %>% 
-  add_recipe(ridge_lasso_recipe) %>% 
+  add_recipe(pokemon_recipe) %>% 
   add_model(ridge_spec)
@@
 
-tune_res <- tune_grid(
+ridge_tune <- tune_grid(
   ridge_workflow,
@@
 )
+
+write_rds(ridge_tune, file = "ridge.rds")
 ```
 ```{r}
-autoplot(tune_res)
+ridge_tune <- read_rds(file = "ridge.rds")
+
+autoplot(ridge_tune)
 ```
@@
 ```{r}
-best_penalty <- select_best(tune_res, metric = 'rmse')
+best_penalty <- select_best(ridge_tune, metric = 'rmse')
 
@@
 ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
+
+write_rds(ridge_final_fit, file = "ridge_final.rds")
+
+ridge_final_fit <- read_rds(file = "ridge_final.rds")
+
 collect_metrics(ridge_final_fit)
@@
 
-tune_res2 <- tune_grid(
+lasso_tune <- tune_grid(
   lasso_workflow,
@@
 )
+
+write_rds(lasso_tune, file = "lasso.rds")
+
+lasso_tune <- read_rds(file = "lasso.rds")
 ```
@@
 ```{r}
-autoplot(tune_res2)
+autoplot(lasso_tune)
 ```
@@
 lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
+
+write_rds(lasso_final_fit, file = "lasso_final.rds")
+
+lasso_final_fit <- read_rds(file = "lasso_final.rds")
+
 ```
@@
 poly_rec <- pokemon_recipe %>% 
-  step_poly(attack, defense, hp, sp_attack, sp_defense, speed, degree = tune())
+  step_poly(base_total, base_happiness, experience_growth, height_m, percentage_male, weight_kg, degree = tune())
 
@@
 
-lm_spec <- linear_reg() %>% 
+poly_spec <- linear_reg() %>% 
   set_mode("regression") %>% 
@@
 poly_wf <- workflow() %>% 
-  add_model(lm_spec) %>% 
+  add_model(poly_spec) %>% 
   add_recipe(poly_rec)
@@
 
-tune_res3 <- tune_grid(
+poly_tuned <- tune_grid(
   poly_wf,
@@
 )
-autoplot(tune_res3)
+
+write_rds(poly_tuned, file = "poly.rds")
+poly_tuned <- read_rds(file = "poly.rds")
+
+autoplot(poly_tuned)
 ```
@@
 ```{r}
-best_poly<- select_best(tune_res3, metric = "rmse")
+best_poly<- select_best(poly_tuned, metric = "rmse")
 poly_final <- finalize_workflow(poly_wf, best_poly)
 poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
+
+write_rds(poly_final_fit, file = "poly_final.rds")
+
+poly_final_fit <- read_rds(file = "poly_final.rds")
 
@@
 ### KNN  
-```{r}
+```{r, warning=FALSE}
 knn_model <- nearest_neighbor(
@@
 
-knn_parameters <- parameters(knn_model)
-
-knn_grd <- grid_regular(knn_parameters, levels = 5)
+knn_grid <- grid_regular(neighbors(range = c(1,15)), levels = 5)
 
@@
     resamples = pokemon_folds,
-    grid = knn_grd
+    grid = knn_grid
   )
+
+write_rds(knn_tune, file = "knn.rds")
+
+knn_tune <- read_rds(file = "knn.rds")
+
 ```
@@
 knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
+
+write_rds(knn_final_fit, file = "knn_final.rds")
+
+knn_final_fit <- read_rds(file = "knn_final.rds")
 ```
@@
 )
+
+write_rds(elastic_tune, file = "elastic.rds")
 ```
@@
 ```{r}
+elastic_tune <- read_rds(file = "elastic.rds")
+
 autoplot(elastic_tune)
+```
+
+```{r}
+# Finding roc_auc of best performing random forest model on folds
+elastic_best_rmse <- elastic_tune %>% 
+  collect_metrics() %>% 
+  arrange(mean)  %>% 
+  slice(73)
+
+elastic_best_rmse
 ```
@@
 elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
+
+write_rds(elastic_final_fit, file = "elastic_final.rds")
+
+elastic_final_fit <- read_rds(file = "elastic_final.rds")
 
@@
 # Creating a regular grid with 8 levels each
-rf_parameter_grid <- grid_regular(mtry(range = c(1, 19)), trees(range = c(200,1000)), min_n(range = c(5,20)), levels = 8)
+rf_parameter_grid <- grid_regular(mtry(range = c(1, 14)), trees(range = c(200,1000)), min_n(range = c(5,20)), levels = 8)
 ```
@@
 # Finding roc_auc of best performing random forest model on folds
-rf_best_roc_auc <- rf_tuned %>% 
+rf_best_rmse <- rf_tuned %>% 
   collect_metrics() %>% 
-  arrange(mean)
+  arrange(mean) %>% 
+  slice(513)
 
-rf_best_roc_auc
+rf_best_rmse
 ```
@@
 rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
-rf_final_fit <- fit(rf_final_workflow, data = pokemon_train)
+rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
 
-rf_final_fit %>% 
+write_rds(rf_final_fit, file = "rf_final.rds")
+
+rf_final_fit <- read_rds(file = "rf_final.rds")
+
+collect_metrics(rf_final_fit) %>% slice(1)
+```
+
+```{r}
+best_rf2 <- select_best(rf_tuned)
+rf_final_workflow2 <- finalize_workflow(rf_workflow, best_rf)
+rf_final_fit2 <- fit(rf_final_workflow, data = pokemon_train)
+
+rf_final_fit2 %>% 
   extract_fit_engine() %>% 
@@
 # Creating regular grid
-boosted_grid <- grid_regular(trees(range = c(10, 2000)), levels = 10)
+boosted_grid <- grid_regular(trees(range = c(5, 200)), learn_rate(range = c(0.01,0.1), trans = identity_trans()), min_n(range = c(40, 60)), levels = 5)
 ```
@@
 # Printing autoplot of results
-autoplot(boosted_tuned)
+autoplot(boosted_tuned, metric = 'rmse')
 ```
@@
 # Finding roc_auc of best performing boosted tree model on folds
-boosted_best_roc_auc <- boosted_tuned %>% 
+boosted_best_rmse <- boosted_tuned %>% 
   collect_metrics() %>% 
-  arrange((mean))
+  arrange(mean) %>% 
+  slice(126)
 
-boosted_best_roc_auc
+boosted_best_rmse
 ```
+
+```{r, eval = FALSE, warning = FALSE}
+# Elastic
+elastic_tuned <- read_rds(file = "elastic_tune.rds")
+
+# Fitting best random forest model to the folds
+best_elastic <- select_best(elastic_tuned)
+elastic_final_workflow <- finalize_workflow(elastic_workflow, best_elastic)
+elastic_final_fit <- fit_resamples(elastic_final_workflow, resamples = pokemon_folds)
+
+# Save
+write_rds(rf_final_fit, file = "elastic_final.rds")
+
+# Polynomial Regression
+poly_tuned <- read_rds(file = "poly_tuned.rds")
+
+# Fitting best random forest model to the folds
+best_poly <- select_best(poly_tuned)
+poly_final_workflow <- finalize_workflow(poly_wf, best_poly)
+poly_final_fit <- fit_resamples(poly_final_workflow, resamples = pokemon_folds)
+
+# Save
+write_rds(rf_final_fit, file = "poly_final.rds")
+
+# Random Forest
+rf_tuned <- read_rds(file = "rf.rds")
+
+# Fitting best random forest model to the folds
+best_rf <- select_best(rf_tuned)
+rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
+rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
+
+# Save
+write_rds(rf_final_fit, file = "rf_final.rds")
+
+# Boosting
+boosted_tuned <- read_rds(file = "boosted.rds")
+
+# Fitting best boosted trees model to the folds
+best_boosted <- select_best(boosted_tuned)
+boosted_final_workflow <- finalize_workflow(boosted_workflow, best_boosted)
+boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
+
+# Save
+write_rds(boosted_final_fit, file = "boosted_final.rds")
+```
+
+```{r}
+
+```

### 2022-12-10 Final_Draft.Rmd
commit

@@
 
-## Introduction
+# Introduction
 
@@
 
-### What Are Pokemon?
+## What Are Pokemon?
 
@@
 
-### What Are We Trying to Do?
+## What Are We Trying to Do?
 
@@
 
-### Why?
+## Why?
 
@@
 
-## Loading Packages and Data
+# Loading Packages and Data
+
+First, let's load in all of our packages and the raw Pokemon data.
 
@@
 library(ranger)
+library(tidytext)
 
@@
 
-## Exploring and Cleaning the Raw Data
+# Exploring and Tidying the Raw Data
+
+## Variable Selection
 
@@
 ```{r}
+# Checking how many unique classifications there are
 pokemon %>%
@@
 
-Now, let's filter out all those unwanted variables!!
+Now, let's filter out all those unwanted variables and our dataset will be ready for further tidying!
 
@@
   clean_names()
-head(pokemon)
 ```
+
+## Tidying the Outcome
 
@@
 # Finding which Pokemon has a character capture rate
-filter(pokemon, capture_rate %in% c("30 (Meteorite)255 (Core)"))
+pokemon %>% 
+  filter(capture_rate %in% c("30 (Meteorite)255 (Core)")) %>% 
+  select(c(name, capture_rate))
 ```
@@
 pokemon$capture_rate <- as.integer(pokemon$capture_rate)
+
+# Displaying the new row
+pokemon %>% 
+  filter(name %in% c("Minior")) %>% 
+  select(c(name, capture_rate))
 ```
 
-Now our outcome is an integer we can use in regression!
+Now the capture rate for Minior is fixed and our outcome is an integer we can use in regression!
 
-### Sources
+## Tidying the Abilities Variable
+
+One last thing we need to do before our data is tidy the `abilities` variable because we currently have a column made up of lists of each Pokemon's abilities.
+However, it is not very useful to analyze 801 unique lists of abilities.
+It might be better to prioritize the abilities that are the most common among the Pokemon.
+Let's look at the top 5 most common abilities and their capture rates.
+
+```{r, warning = FALSE}
+# Creating a copy of the dataset to work with
+pokemon_copy <- pokemon
+
+# Splitting up the lists in the abilities column
+pokemon_copy$abilities <- pokemon_copy$abilities %>% 
+  str_split(pattern = "', '") %>% 
+  map(~ str_replace_all(string = ., pattern = "[[:punct:]]", ""))
+
+# Unnesting so the abilities are separate rows
+pokemon_copy2 <- pokemon_copy %>% 
+  unnest()
+
+# Viewing the top 5 abilities by count and their capture rates
+pokemon_copy2 %>% 
+  group_by(abilities) %>%  
+  summarise(count = n(), capture_rate = mean(capture_rate)) %>% 
+  arrange(desc(count)) %>% 
+  head(5)
+```
+
+The top 5 abilities appear to have varying capture rates that are not too close to one another.
+So, it might prove meaningful to include whether a Pokemon has one of the top 5 most common abilities or not.
+Now, we must add in 5 more variables indicating whether a Pokemon has each of these abilities or not.
+
+```{r}
+# Adding in 5 more variables indicating whether a Pokemon has each of the top 5 abilities or not
+pokemon <- pokemon %>% 
+  mutate(has_sturdy = if_else(grepl("Sturdy", pokemon$abilities), 1, 0)) %>%
+  mutate(has_swift_swim = if_else(grepl("Swift Swim", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_keen_eye = if_else(grepl("Keen Eye", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_chlorophyll = if_else(grepl("Chlorophyll", pokemon$abilities), 1, 0)) %>% 
+  mutate(has_levitate = if_else(grepl("Levitate", pokemon$abilities), 1, 0))
+
+# Creating a final dataset csv
+write_csv(pokemon, "pokemon_final.csv")
+```
+
+Our dataset is now tidied and ready for some exploratory data analysis!
+
+# Exploratory Data Analysis
+
+## Loading Data and Converting Factors
+
+First, we need to read and load in the data and convert any categorical variables to factors.
+
+```{r, message = FALSE}
+# Reading in the data
+pokemon_data <- read_csv("pokemon_final.csv")
+
+# Changing categorical variables to factors
+pokemon_data$type1 <- as.factor(pokemon_data$type1)
+pokemon_data$generation <- as.factor(pokemon_data$generation)
+pokemon_data$is_legendary <- as.factor(pokemon_data$is_legendary)
+pokemon_data$has_sturdy <- as.factor(pokemon_data$has_sturdy)
+pokemon_data$has_swift_swim <- as.factor(pokemon_data$has_swift_swim)
+pokemon_data$has_keen_eye <- as.factor(pokemon_data$has_keen_eye)
+pokemon_data$has_chlorophyll <- as.factor(pokemon_data$has_chlorophyll)
+pokemon_data$has_levitate <- as.factor(pokemon_data$has_levitate)
+
+# Loading the data
+head(pokemon_data)
+```
+
+## Missing Data
+
+Before we move on to exploring the variables in our data, we must check for any missing data because that could potentially cause issues.
+
+```{r}
+pokemon_data %>% 
+  summary()
+```
+
+We can see that `height_m`, `percentage_male`, and `weight_kg` all have missing values: 20, 98, and 20, respectively.
+It appears that the variables with missing data relate more so to the physical characteristics of the Pokemon and not as much to the capabilities of the Pokemon.
+Because these are not very high proportions of missing data, we can deal with this issue by imputing the missing values using a linear regression of our other variables.
+This will be done later in the recipe creation!
+
+## Visual EDA
+
+### Capture Rate
+
+First, let's explore the distribution of our outcome, `capture_rate`.
+
+```{r}
+# Distribution of capture rate
+ggplot(pokemon_data, aes(capture_rate)) +
+  geom_bar(fill='red3')
+```
+
+The possible values of `capture_rate` range from 0 to 255 and the `capture_rate` of the Pokemon in our data appears to range from 3 to 255.
+There is also a peak at about 45, meaning the most common capture rate for a Pokemon in our data is 45.
+
+### Generation
+
+The data contains 7 different generations of Pokemon.
+The generations contain Pokemon from each of the 7 generations of Pokemon games.
+This poses the question of whether there are specific generations that have higher capture rates overall than any other generations.
+
+```{r}
+# Boxplot of generation vs. capture rate
+ggplot(pokemon, aes(x=generation, y=capture_rate, group=generation)) +
+  geom_boxplot(aes(fill = generation))
+```
+
+From this plot, we can see that the medians, maxima, minima, and first quartiles of the capture rates for each of the generations are the same or quite close to one another.
+However, the third quartile of generations 2, 5, and 7 appear to be less than the other generations.
+This does not seem as if it will make much different as the rest of the overall distribution is quite similar between generations.
+So, it appears that `capture_rate` does not vary much depending on specifically `generation` alone.
+
+### Legendary Status
+
+Legendary Pokemon are especially powerful Pokemon that players rarely encounter.
+From this definition, it seems likely that it would be harder to capture a Pokemon that is legendary than one that is not.
+
+```{r}
+ggplot(pokemon_data, aes(capture_rate)) + 
+  geom_bar(aes(fill = is_legendary)) +
+  scale_fill_manual(values = c('red2', 'blue4'))
+```
+
+From the graph, we can see that, although there are still non-legendary Pokemon in the lower end of the capture rate, there are no legendary Pokemon with a capture rate above 50.
+Also, the lowest capture rate in our data, 3, is made up of almost all legendary Pokemon.
+It seems that legendary Pokemon have a lower capture rate than non-legendary Pokemon, which is what was expected.
+
+### Correlation Plot
+
+Now, let's explore the overall relationships between all of the continuous, non-missing variables using a correlation plot.
+
+```{r}
+# Selecting only numeric values
+pokemon_numeric <- pokemon_data %>%
+  select_if(is.numeric)
+# Taking out variables with missing values as they will return NA
+pokemon_numeric <- pokemon_numeric[, !names(pokemon_numeric) %in% c("height_m", "weight_kg", "percentage_male")]
+
+# Correlation matrix
+pokemon_cor <- cor(pokemon_numeric)
+
+# Visualization of correlation matrix
+pokemon_corrplot <- corrplot(pokemon_cor, method = "number")
+```
+
+One relationship that stands out is the relatively strong negative correlation between `base_total` and `capture_rate` (-0.71), which indicates that it may be useful to use `base_total` in our analysis rather than the six individual battle statistics (we cannot use all of them because the battle statistics add up to the exact `base_total`).
+This relationship implies that, as a Pokemon has a higher total of battle statistics, it becomes more difficult to capture them (lower `capture_rate`).
+
+Another apparent relationship is that all six battle statistics are moderately positively correlated with `base_total` and moderately negatively correlated with `capture_rate`.
+This implies that, as the battle statistics increase, so does their total, which makes sense as they add up to the `base_total`.
+Also, as the six battle statistics increase, the Pokemon becomes more difficult to capture (lower `capture_rate`).
+Since the battle statistics have a similar relationship with `capture_rate` as `base_total`, this further justifies using only `base_total` rather than the six individual battle statistics.
+
+# Setting up for the Models
+
+It's time to start fitting models to our data to see if we really can predict the capture rate of a Pokemon based on the predictors we have.
+However, we first have to set up our data by splitting it, creating the recipe, and creating folds for k-fold cross validation.
+
+## Data Split
+
+The first step we have to take before fitting any models is splitting our data into a training and testing set.
+The training set will be used to do just that: train our models.
+The testing set acts as a test in the sense that our models will not be able to train on that data.
+So, once we fit whichever model we decide is the "best" (usually based on lowest RMSE or root mean squared error for regression) to our testing set, we will see how it truly performs on new data.
+By splitting our data into testing and training sets, we avoid over-fitting because the model is not using all of the available data to learn.
+The split I have chosen is a 70/30 split, which means 70% of the data will go towards the training set and 30% of the data will go towards the testing set.
+This way, most of our data is being used to train the model; however, we still have an adequate amount of data to test the model on.
+Also, the split is stratified on the outcome variable, `capture_rate`, to ensure that both the training and the testing data have an equal distribution of `capture_rate`.
+
+```{r}
+# Setting the seed for reproducibility
+set.seed(0124)
+
+# Splitting the data (70/30 split, stratify on capture rate)
+pokemon_split <- initial_split(pokemon_data, prop = 0.7, strata = capture_rate)
+pokemon_train <- training(pokemon_split)
+pokemon_test <- testing(pokemon_split)
+```
+
+Before we move on, we should verify that the data split correctly.
+
+```{r}
+nrow(pokemon_train)/nrow(pokemon_data)
+nrow(pokemon_test)/nrow(pokemon_data)
+```
+
+The training set has about 70% of the data and the testing set has about 30% of the data. So, the data was split correctly between the training and testing sets.
+
+## Recipe Creation
+
+Throughout our analysis, we will be using essentially the same predictors, conditions, and outcome.
+So, we will create one universal recipe to use for all of our models (and slightly adjust this recipe if needed).
+Each model will take this recipe and work with it using the methods associated with that model.
+
+We will be using the same 14 predictors: `base_total`, `base_happiness`, `experience_growth`, `height_m`, `percentage_male`, `weight_kg`, `type1`, `generation`, `is_legendary`, `has_sturdy`, `has_swift_swim`, `has_keen_eye`, `has_chlorophyll`, and `has_levitate`.
+As mentioned in the exploratory data analysis, it would be best to use `base_total` instead of the individual six battle statistics because it has a higher negative correlation with the outcome and also acts as a summary of each Pokemon's six battle statistics.
+
+In order to deal with missingness, we will impute the missing data in `percentage_male`, `height_m`, and `weight_kg` using a linear regression of all of the other predictors in the recipe.
+
+Since `type1` has 18 types, we will prioritize the types made up of the most Pokemon and group the rest of the Pokemon into type 'Other'.
+
+We will make `type1`, `generation`, `is_legendary`, `has_sturdy`, `has_swift_swim`, `has_keen_eye`, `has_chlorophyll`, and `has_levitate` into dummy variables because they are categorical and not continous.
+
+Finally, we normalize our variables by centering and scaling.
+
+```{r}
+# Creating recipe
+pokemon_recipe <- recipe(capture_rate ~ base_total + base_happiness + experience_growth + height_m + percentage_male + weight_kg + type1 + generation + is_legendary + has_sturdy + has_swift_swim + has_keen_eye + has_chlorophyll + has_levitate, data = pokemon_train) %>%
+  # imputing percentage male
+  step_impute_linear(percentage_male, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  # imputing height
+  step_impute_linear(height_m, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  # imputing weight
+  step_impute_linear(weight_kg, impute_with = imp_vars(base_total, base_happiness, experience_growth, type1, generation, is_legendary, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate)) %>%
+  # Only prioritizing top types of Pokemon
+  step_other(type1) %>% 
+  # dummy coding nominal variables
+  step_dummy(generation, is_legendary, type1, has_sturdy, has_swift_swim, has_keen_eye, has_chlorophyll, has_levitate) %>% 
+  # normalizing
+  step_center(all_predictors()) %>% 
+  step_scale(all_predictors())
+```
+
+## K-Fold Cross Validation
+
+We will use k-fold stratified cross validation in order to train the data on 10 mini datasets (or folds) created using the training data.
+Since we will be essentially using 10 mini training/testing sets to train our data, this will give us a better estimate of each model's true performance.
+We stratify on the outcome, `capture_rate`, to make sure the data in each fold is not imbalanced.
+
+```{r}
+# Creating folds
+pokemon_folds <- vfold_cv(pokemon_train, v = 10, strata = capture_rate)
+```
+
+# Model Building
+
+It is now time to build our models!
+Since the models take a very long time to run, the results from each of the models has been saved to avoid rerunning the models every time.
+I have chosen Root Mean Squared Error (RMSE) as my metric because it works as an overall metric for all models.
+The RMSE is one of the most commonly used measures for evaluating the performance of regression models by showing how far the model's predictions are from the true values using Euclidian distance.
+Since RMSE measures distance, it is important that we normalize our data, which we did in the recipe.
+Also, I have fit 8 models to the Pokemon data; however, we will only be conducting further analysis on the 4 best-performing models.
+Let's get to building our models!
+
+## Fitting the models
+
+Each of the models had a very similar process. I will detail it below and include the code for each of the models under that step (however, the code will not be evaluated here to save time).
+
+For each of the models, you must conduct these steps to fit them:
+
+1. Set up the model by specifying the model you wish to fit, the parameters you want to tune, the engine the model comes from, and the mode (regression or classification) if necessary.
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# LINEAR REGRESSION 
+lm_model <- linear_reg() %>% 
+  set_engine("lm")
+
+# RIDGE REGRESSION
+# Tuning penalty and setting mixture to 0 to specify ridge
+ridge_spec <- linear_reg(mixture = 0, 
+                         penalty = tune()) %>% 
+  set_mode("regression") %>% 
+  set_engine("glmnet")
+
+# LASSO REGRESSION
+# Tuning penalty and setting mixture to 1 to specify lasso
+lasso_spec <- linear_reg(penalty = tune(), 
+                         mixture = 1) %>% 
+  set_mode("regression") %>% 
+  set_engine("glmnet")
+
+# POLYNOMIAL REGRESSION
+# Adjusting the recipe because the tuning parameter must be added in the recipe for polynomial regression
+# Tuning the degree
+poly_rec <- pokemon_recipe %>% 
+  step_poly(base_total, base_happiness, experience_growth, height_m, percentage_male, weight_kg, degree = tune())
+
+poly_spec <- linear_reg() %>% 
+  set_mode("regression") %>% 
+  set_engine("lm")
+
+# K NEAREST NEIGHBORS
+# Tuning the number of neighbors
+knn_model <- nearest_neighbor(neighbors = tune()) %>% 
+  set_mode("regression") %>% 
+  set_engine("kknn")
+
+# ELASTIC NET
+# Tuning penalty and mixture
+elastic_spec <- linear_reg(penalty = tune(), 
+                           mixture = tune()) %>% 
+  set_mode("regression") %>% 
+  set_engine("glmnet")
+
+# RANDOM FOREST
+# Tuning mtry (number of predictors), trees, and min_n (number of minimum values in each node)
+rf_spec <- rand_forest(mtry = tune(), 
+                       trees = tune(), 
+                       min_n = tune()) %>% 
+  set_engine("ranger", importance = "impurity") %>% 
+  set_mode("regression")
+
+# BOOSTED TREES
+# Tuning trees, learn_rate (the learning rate), and min_n
+boosted_spec <- boost_tree(trees = tune(),
+                           learn_rate = tune(),
+                           min_n = tune()) %>%
+  set_engine("xgboost") %>%
+  set_mode("regression")
+```
+
+2. Set up the workflow for the model and add the model and the recipe.
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# LINEAR REGRESSION 
+lm_workflow <- workflow() %>% 
+  add_model(lm_model) %>% 
+  add_recipe(pokemon_recipe)
+
+# RIDGE REGRESSION
+ridge_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(ridge_spec)
+
+# LASSO REGRESSION
+lasso_workflow <- workflow() %>% 
+  add_recipe(ridge_lasso_recipe) %>% 
+  add_model(lasso_spec)
+
+# POLYNOMIAL REGRESSION
+poly_wf <- workflow() %>% 
+  add_model(poly_spec) %>% 
+  add_recipe(poly_rec)
+
+# K NEAREST NEIGHBORS
+knn_workflow <- workflow() %>% 
+  add_model(knn_model) %>% 
+  add_recipe(pokemon_recipe)
+
+# ELASTIC NET
+elastic_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(elastic_spec)
+
+# RANDOM FOREST
+rf_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(rf_spec)
+
+# BOOSTED TREES
+boosted_workflow <- workflow() %>% 
+  add_recipe(pokemon_recipe) %>% 
+  add_model(boosted_spec)
+```
+
+3. Create a tuning grid to specify the ranges of the parameters you wish to tune as well as how many levels of each.
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# LINEAR REGRESSION 
+# No grid because no tuning parameters
+
+# RIDGE REGRESSION
+penalty_grid <- grid_regular(penalty(range = c(-5,5)), levels = 50)
+
+# LASSO REGRESSION
+# Same grid as ridge
+
+# POLYNOMIAL REGRESSION
+degree_grid <- grid_regular(degree(range = c(1,5)), levels = 5)
+
+# K NEAREST NEIGHBORS
+knn_grid <- grid_regular(neighbors(range = c(1,15)), levels = 5)
+
+# ELASTIC NET
+elastic_grid <- grid_regular(penalty(range = c(-5, 5)), mixture(range = c(0,1)), levels = 10)
+
+# RANDOM FOREST
+rf_parameter_grid <- grid_regular(mtry(range = c(1, 14)), trees(range = c(200,1000)), min_n(range = c(5,20)), levels = 8)
+
+# BOOSTED TREES
+boosted_grid <- grid_regular(trees(range = c(5, 200)), learn_rate(range = c(0.01,0.1), trans = identity_trans()), min_n(range = c(40, 60)), levels = 5)
+```
+
+4. Tune the model and specify the workflow, k-fold cross validation folds, and the tuning grid for our chosen parameters to tune.
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# LINEAR REGRESSION 
+# No tuning
+
+# RIDGE REGRESSION
+ridge_tune <- tune_grid(
+  ridge_workflow,
+  resamples = pokemon_folds,
+  grid = penalty_grid
+)
+
+# LASSO REGRESSION
+lasso_tune <- tune_grid(
+  lasso_workflow,
+  resamples = pokemon_folds,
+  grid = penalty_grid
+)
+
+# POLYNOMIAL REGRESSION
+poly_tuned <- tune_grid(
+  poly_wf,
+  resamples = pokemon_folds,
+  grid = degree_grid
+)
+
+# K NEAREST NEIGHBORS
+knn_tune <- tune_grid(
+    knn_workflow,
+    resamples = pokemon_folds,
+    grid = knn_grid
+)
+
+# ELASTIC NET
+elastic_tune <- tune_grid(
+  elastic_workflow,
+  resamples = pokemon_folds,
+  grid = elastic_grid
+)
+
+# RANDOM FOREST
+rf_tune_res <- tune_grid(
+  rf_workflow,
+  resamples = pokemon_folds,
+  grid = rf_parameter_grid
+)
+
+# BOOSTED TREES
+boosted_tune_res <- tune_grid(
+  boosted_workflow,
+  resamples = pokemon_folds,
+  grid = boosted_grid
+)
+```
+
+5. Save the tuned models to an RDS file to avoid rerunning the model.
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# write_rds() to save
+
+# LINEAR REGRESSION 
+# No tuning
+
+# RIDGE REGRESSION
+write_rds(ridge_tune, file = "ridge.rds")
+
+# LASSO REGRESSION
+write_rds(lasso_tune, file = "lasso.rds")
+
+# POLYNOMIAL REGRESSION
+write_rds(poly_tuned, file = "poly.rds")
+
+# K NEAREST NEIGHBORS
+write_rds(knn_tune, file = "knn.rds")
+
+# ELASTIC NET
+write_rds(elastic_tune, file = "elastic.rds")
+
+# RANDOM FOREST
+write_rds(rf_tune_res, file = "rf.rds")
+
+# BOOSTED TREES
+write_rds(boosted_tune_res, file = "boosted.rds")
+```
+
+6. Load back in the saved files.
+
+```{r, warning=FALSE, message=FALSE}
+# read_rds() to load back in
+
+# LINEAR REGRESSION 
+# No tuning
+
+# RIDGE REGRESSION
+ridge_tuned <- read_rds(file = "ridge.rds")
+
+# LASSO REGRESSION
+lasso_tuned <- read_rds(file = "lasso.rds")
+
+# POLYNOMIAL REGRESSION
+poly_tuned <- read_rds(file = "poly.rds")
+
+# K NEAREST NEIGHBORS
+knn_tuned <- read_rds(file = "knn.rds")
+
+# ELASTIC NET
+elastic_tuned <- read_rds(file = "elastic.rds")
+
+# RANDOM FOREST
+rf_tuned <- read_rds(file = "rf.rds")
+
+# BOOSTED TREES
+boosted_tuned <- read_rds(file = "boosted.rds")
+```
+
+7.
+Select the best tuned parameters for the model (based on RMSE), adjust the model workflow to include that version of the model, and fit the model with the best tuned parameters to the folds.
+Make sure to save the final fit.
+
+```{r, eval=FALSE, warning=FALSE, message=FALSE}
+# select_best() to select the best tuned parameters for the model (based on RMSE)
+# finalize_workflow() to adjust the model workflow to include that version of the model
+# fit_resamples() to fit the model with the best tuned parameter to the folds
+
+# LINEAR REGRESSION 
+# No need to select the best fit as linear regression had no tuning -> we just fit the one linear regression model we fit
+lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
+write_rds(lm_fit, file = "lm_final.rds")
+
+# RIDGE REGRESSION
+best_penalty <- select_best(ridge_tune, metric = 'rmse')
+ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
+ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
+write_rds(ridge_final_fit, file = "ridge_final.rds")
+
+# LASSO REGRESSION
+best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
+lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
+lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
+write_rds(lasso_final_fit, file = "lasso_final.rds")
+
+# POLYNOMIAL REGRESSION
+best_poly<- select_best(poly_tuned, metric = "rmse")
+poly_final <- finalize_workflow(poly_wf, best_poly)
+poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
+write_rds(poly_final_fit, file = "poly_final.rds")
+
+# K NEAREST NEIGHBORS
+best_knn <- select_best(knn_tune, metric = "rmse")
+knn_final <- finalize_workflow(knn_workflow, best_knn)
+knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
+write_rds(knn_final_fit, file = "knn_final.rds")
+
+# ELASTIC NET
+best_elastic <- select_best(elastic_tune, metric = "rmse")
+elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
+elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
+write_rds(elastic_final_fit, file = "elastic_final.rds")
+
+# RANDOM FOREST
+best_rf <- select_best(rf_tuned)
+rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
+rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
+write_rds(rf_final_fit, file = "rf_final.rds")
+
+# BOOSTED TREES
+best_boosted <- select_best(boosted_tuned)
+boosted_final_workflow <- finalize_workflow(boosted_workflow, best_boosted)
+boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
+write_rds(boosted_final_fit, file = "boosted_final.rds")
+```
+
+8. Load in each of the best models and collect the RMSE's of each of the best model fits for comparison with one another.
+
+```{r, warning=FALSE, message=FALSE}
+# collect_metrics() to collect the RMSE
+# slice() to only collect the RMSE and not R squared
+
+# LINEAR REGRESSION 
+lm_final_fit <- read_rds(file = "lm_final.rds")
+lm_rmse <- collect_metrics(lm_final_fit) %>% slice(1)
+
+# RIDGE REGRESSION
+ridge_final_fit <- read_rds(file = "ridge_final.rds")
+ridge_rmse <- collect_metrics(ridge_final_fit) %>% slice(1)
+
+# LASSO REGRESSION
+lasso_final_fit <- read_rds(file = "lasso_final.rds")
+lasso_rmse <- collect_metrics(lasso_final_fit) %>% slice(1)
+
+# POLYNOMIAL REGRESSION
+poly_final_fit <- read_rds(file = "poly_final.rds")
+poly_rmse <- collect_metrics(poly_final_fit) %>% slice(1)
+
+# K NEAREST NEIGHBORS
+knn_final_fit <- read_rds(file = "knn_final.rds")
+knn_rmse <- collect_metrics(knn_final_fit) %>% slice(1)
+
+# ELASTIC NET
+elastic_final_fit <- read_rds(file = "elastic_final.rds")
+elastic_rmse <- collect_metrics(elastic_final_fit) %>% slice(1)
+
+# RANDOM FOREST
+rf_final_fit <- read_rds(file = "rf_final.rds")
+rf_rmse <- collect_metrics(rf_final_fit) %>% slice(1)
+
+# BOOSTED TREES
+boosted_final_fit <- read_rds(file = "boosted_final.rds")
+boosted_rmse <- collect_metrics(boosted_final_fit) %>% slice(1)
+```
+
+# Model Results
+
+It's finally time to compare the results of all of our models and see which ones performed the best!
+
+```{r}
+# Creating a tibble of all the models and their RMSE
+final_compare_tibble <- tibble(Model = c("Linear Regression", "Ridge Regression", "Lasso Regression", "Polynomial Regression", "K Nearest Neighbors", "Elastic Net", "Random Forest", "Boosted Trees"), RMSE = c(lm_rmse$mean, ridge_rmse$mean, lasso_rmse$mean, poly_rmse$mean, knn_rmse$mean, elastic_rmse$mean, rf_rmse$mean, boosted_rmse$mean))
+
+# Arranging by lowest RMSE
+final_compare_tibble <- final_compare_tibble %>% 
+  arrange(RMSE)
+
+final_compare_tibble
+```
+
+From the performance of the models on the cross-validation data, we can see that the random forest performed the best!
+The other three models we will focus on as well are boosted trees, polynomial regression, and elastic net because those placed right after the random forest.
+One key thing to note from these results is that it appears the linear and simpler models did the worst, which indicates that our data is probably nonlinear.
+
+## Model Autoplots
+
+The autoplot in R allows us to visualize the effect of each of the tuned parameters on the performance of each model.
+In the autoplots, the performance of the models is measured by the RMSE (the lower the RMSE, the better the model has performed).
+
+### Elastic Net Plot
+
+```{r}
+autoplot(elastic_tuned, metric = 'rmse')
+```
+
+For the elastic net, we tuned penalty and mixture at 10 different levels.
+From the plot, it seems that lower to mid-range penalty values appear to do the best.
+As the penalty term becomes too large, the model does worse because the coefficients of the predictors are being reduced to very small values (due to the penalty), which makes it much more difficult for the model to predict.
+Interestingly, although the RMSE appears to be lowest at a penalty value slightly greater than one (which matches the overall trend), the lowest RMSE appears to have a high mixture value.
+It seems that lower values of mixture only perform better than higher values of mixture when the penalty term is very high.
+Otherwise, when the penalty term is low to mid-range, higher values of mixture appear to perform better.
+This lines up with our overall model results as the lasso regression performed closer to the elastic net than the ridge regression (since lasso regression has a mixture value of 1 and ridge regression has a mixture value of 0).
+
+### Polynomial Regression Plot
+
+```{r}
+autoplot(poly_tuned, metric = 'rmse')
+```
+
+For the polynomial regression, we tuned the degree of all continuous variables at 5 different levels.
+As we can see from the plot, the model performs better when a degree is added to the predictors (except when the degree gets too high).
+This further affirms our initial idea that the data is nonlinear.
+From the plot, it appears that the model does the best at a polynomial degree of 3 and does substantially worse after a polynomial degree of 4.
+
+### Boosted Trees Autoplot
+
+```{r}
+autoplot(boosted_tuned, metric = 'rmse')
+```
+
+For the boosted trees, we tuned the learning rate, number of trees, and minimal node size with 5 different levels.
+It may be useful to note that some tweaking of the model was done prior to finalizing the ranges for these parameters.
+The model did substantially worse when the minimal node size was a lower number, which is why the range 40 to 60 was chosen.
+Also, the model did not do as well when the learning rate was too high.
+A high learning rate causes the model to learn faster, but it also trains the model less and makes it less generalized.
+So, the range 0.01 to 0.1 was chosen to see which smaller learning rate value performs the best.
+From the starting points of each of the plots, it appears that the model does better at a higher learning rate.
+This means that the model does better when it is learning faster.
+Also, it appears that generally, once there are more than ~50 trees, the number of trees does not make much of a change on the performance of the model.
+However, the model appears to do worse when there are less than ~50 trees.
+The minimal node size does not seem to have much of an effect on the performance of the model.
+
+### Random Forest Autoplot
+
+```{r}
+autoplot(rf_tuned, metric = 'rmse')
+```
+
+For the random forest,
+
+# Results of the Best Model
+
+```{r, eval = FALSE}
+# Fitting to the training data
+best_rf_train <- select_best(rf_tuned)
+rf_final_workflow_train <- finalize_workflow(rf_workflow, best_rf)
+rf_final_fit_train <- fit(rf_final_workflow, data = pokemon_train)
+
+write_rds(rf_final_fit, file = "rf_final_train.rds")
+```
+
+```{r}
+rf_final_fit_train <- read_rds(file = "rf_final_train.rds")
+
+# VIP plot
+rf_final_fit_train %>% 
+  extract_fit_engine() %>% 
+  vip()
+```
+
+```{r}
+# Indicating the desired metric
+pokemon_metric <- metric_set(rmse)
+
+# Creating the predicted vs. actual value tibble
+pokemon_tibble <- predict(rf_final_fit, new_data = pokemon_test %>% select(-capture_rate))
+pokemon_tibble <- bind_cols(pokemon_tibble, pokemon_test %>% select(capture_rate))
+
+# Collecting the rmse of the model on the testing data
+pokemon_tibble_metrics <- pokemon_metric(pokemon_tibble, truth = capture_rate, estimate = .pred)
+pokemon_tibble_metrics
+```
+
+# Sources

### 2022-12-10 Final_Draft.Rmd
commit

@@
 
-```{r}
+```{r, eval=FALSE}
 # Adding in 5 more variables indicating whether a Pokemon has each of the top 5 abilities or not
@@
 # Creating a final dataset csv
-write_csv(pokemon, "pokemon_final.csv")
+write_csv(pokemon, "data/pokemon_final.csv")
 ```
@@
 # Reading in the data
-pokemon_data <- read_csv("pokemon_final.csv")
+pokemon_data <- read_csv("data/pokemon_final.csv")
 
@@
 # RIDGE REGRESSION
-write_rds(ridge_tune, file = "ridge.rds")
+write_rds(ridge_tune, file = "data/tuned_models/ridge.rds")
 
 # LASSO REGRESSION
-write_rds(lasso_tune, file = "lasso.rds")
+write_rds(lasso_tune, file = "data/tuned_models/lasso.rds")
 
 # POLYNOMIAL REGRESSION
-write_rds(poly_tuned, file = "poly.rds")
+write_rds(poly_tuned, file = "data/tuned_models/poly.rds")
 
 # K NEAREST NEIGHBORS
-write_rds(knn_tune, file = "knn.rds")
+write_rds(knn_tune, file = "data/tuned_models/knn.rds")
 
 # ELASTIC NET
-write_rds(elastic_tune, file = "elastic.rds")
+write_rds(elastic_tune, file = "data/tuned_models/elastic.rds")
 
 # RANDOM FOREST
-write_rds(rf_tune_res, file = "rf.rds")
+write_rds(rf_tune_res, file = "data/tuned_models/rf.rds")
 
 # BOOSTED TREES
-write_rds(boosted_tune_res, file = "boosted.rds")
+write_rds(boosted_tune_res, file = "data/tuned_models/boosted.rds")
 ```
@@
 # RIDGE REGRESSION
-ridge_tuned <- read_rds(file = "ridge.rds")
+ridge_tuned <- read_rds(file = "data/tuned_models/ridge.rds")
 
 # LASSO REGRESSION
-lasso_tuned <- read_rds(file = "lasso.rds")
+lasso_tuned <- read_rds(file = "data/tuned_models/lasso.rds")
 
 # POLYNOMIAL REGRESSION
-poly_tuned <- read_rds(file = "poly.rds")
+poly_tuned <- read_rds(file = "data/tuned_models/poly.rds")
 
 # K NEAREST NEIGHBORS
-knn_tuned <- read_rds(file = "knn.rds")
+knn_tuned <- read_rds(file = "data/tuned_models/knn.rds")
 
 # ELASTIC NET
-elastic_tuned <- read_rds(file = "elastic.rds")
+elastic_tuned <- read_rds(file = "data/tuned_models/elastic.rds")
 
 # RANDOM FOREST
-rf_tuned <- read_rds(file = "rf.rds")
+rf_tuned <- read_rds(file = "data/tuned_models/rf.rds")
 
 # BOOSTED TREES
-boosted_tuned <- read_rds(file = "boosted.rds")
+boosted_tuned <- read_rds(file = "data/tuned_models/boosted.rds")
 ```
@@
 lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
-write_rds(lm_fit, file = "lm_final.rds")
+write_rds(lm_fit, file = "data/final_models/lm_final.rds")
 
@@
 ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
-write_rds(ridge_final_fit, file = "ridge_final.rds")
+write_rds(ridge_final_fit, file = "data/final_models/ridge_final.rds")
 
@@
 lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
-write_rds(lasso_final_fit, file = "lasso_final.rds")
+write_rds(lasso_final_fit, file = "data/final_models/lasso_final.rds")
 
@@
 poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
-write_rds(poly_final_fit, file = "poly_final.rds")
+write_rds(poly_final_fit, file = "data/final_models/poly_final.rds")
 
@@
 knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
-write_rds(knn_final_fit, file = "knn_final.rds")
+write_rds(knn_final_fit, file = "data/final_models/knn_final.rds")
 
@@
 elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
-write_rds(elastic_final_fit, file = "elastic_final.rds")
+write_rds(elastic_final_fit, file = "data/final_models/elastic_final.rds")
 
@@
 rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
-write_rds(rf_final_fit, file = "rf_final.rds")
+write_rds(rf_final_fit, file = "data/final_models/rf_final.rds")
 
@@
 boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
-write_rds(boosted_final_fit, file = "boosted_final.rds")
+write_rds(boosted_final_fit, file = "data/final_models/boosted_final.rds")
 ```
@@
 # LINEAR REGRESSION 
-lm_final_fit <- read_rds(file = "lm_final.rds")
+lm_final_fit <- read_rds(file = "data/final_models/lm_final.rds")
 lm_rmse <- collect_metrics(lm_final_fit) %>% slice(1)
@@
 # RIDGE REGRESSION
-ridge_final_fit <- read_rds(file = "ridge_final.rds")
+ridge_final_fit <- read_rds(file = "data/final_models/ridge_final.rds")
 ridge_rmse <- collect_metrics(ridge_final_fit) %>% slice(1)
@@
 # LASSO REGRESSION
-lasso_final_fit <- read_rds(file = "lasso_final.rds")
+lasso_final_fit <- read_rds(file = "data/final_models/lasso_final.rds")
 lasso_rmse <- collect_metrics(lasso_final_fit) %>% slice(1)
@@
 # POLYNOMIAL REGRESSION
-poly_final_fit <- read_rds(file = "poly_final.rds")
+poly_final_fit <- read_rds(file = "data/final_models/poly_final.rds")
 poly_rmse <- collect_metrics(poly_final_fit) %>% slice(1)
@@
 # K NEAREST NEIGHBORS
-knn_final_fit <- read_rds(file = "knn_final.rds")
+knn_final_fit <- read_rds(file = "data/final_models/knn_final.rds")
 knn_rmse <- collect_metrics(knn_final_fit) %>% slice(1)
@@
 # ELASTIC NET
-elastic_final_fit <- read_rds(file = "elastic_final.rds")
+elastic_final_fit <- read_rds(file = "data/final_models/elastic_final.rds")
 elastic_rmse <- collect_metrics(elastic_final_fit) %>% slice(1)
@@
 # RANDOM FOREST
-rf_final_fit <- read_rds(file = "rf_final.rds")
+rf_final_fit <- read_rds(file = "data/final_models/rf_final.rds")
 rf_rmse <- collect_metrics(rf_final_fit) %>% slice(1)
@@
 # BOOSTED TREES
-boosted_final_fit <- read_rds(file = "boosted_final.rds")
+boosted_final_fit <- read_rds(file = "data/final_models/boosted_final.rds")
 boosted_rmse <- collect_metrics(boosted_final_fit) %>% slice(1)
@@
 
-For the random forest,
+For the random forest, we tuned the the minimal node size, the number of randomly selected predictors, and the number of trees.
+The number of trees does not seem to have much effect on the performance of the model.
+The minimal node size does not appear to have a huge effect on the performance, but the smaller values of `min_n` appear to have slightly lower RMSE values (at least judging by the starting point).
+The optimal node size appears to be at 5 (which lines up with the aforementioned observation).
+The number of predictors appears to have a greatest effect on performance.
+From the plots, it appears that, overall, a greater number of predictors renders a better performance.
+Actually, it seems that including all 14 predictors gives the lowest RMSE here.
+Based on the RMSE values in the plots, this model definitely performs the best!
 
@@
 
-```{r, eval = FALSE}
+## Performance on the Folds
+
+So, the random forest performed the best out of all 8 of our models. But which tuned parameters were chosen as the best random forest model?
+
+```{r}
+# Collecting metrics on the tuned random forest and slicing the lowest RMSE
+rf_tuned %>% 
+  collect_metrics() %>% 
+  arrange(mean) %>% 
+  slice(513)
+```
+
+Random forest #40 with all 14 predictors, 657 trees, and a minimal node size of 5 performed the best with an RMSE of 43.79638!
+
+## Fitting to Training Data
+
+Now, we will take that best model from the tuned random forest and fit it to the training data. This will train that random forest one more time on the entire training data set.
+
+```{r, eval = FALSE, warning = FALSE}
 # Fitting to the training data
@@
 
-write_rds(rf_final_fit, file = "rf_final_train.rds")
+# Save
+write_rds(rf_final_fit, file = "data/rf_final/rf_final_train.rds")
 ```
@@
 ```{r}
-rf_final_fit_train <- read_rds(file = "rf_final_train.rds")
+# Reading in the random forest fit on the training data
+rf_final_fit_train <- read_rds(file = "data/rf_final/rf_final_train.rds")
 
-# VIP plot
+# Collecting the RMSE of the random forest fit on the training data
 rf_final_fit_train %>% 
-  extract_fit_engine() %>% 
-  vip()
+  collect_metrics() %>% 
+  slice(1)
+```
+
+The chosen random forest model had an RMSE of 43.95341 on the training data. This is quite close to the RMSE the model had on the folds.
+
+## Testing the Model
+
+Now, it's time to test our random forest model to see how it performs on data that it has not been trained on at all: the testing data set.
+
+```{r, eval = FALSE}
+# Creating the predicted vs. actual value tibble
+pokemon_tibble <- predict(rf_final_fit_train, new_data = pokemon_test %>% select(-capture_rate))
+pokemon_tibble <- bind_cols(pokemon_tibble, pokemon_test %>% select(capture_rate))
+
+# Save
+write_rds(pokemon_tibble, file = "data/rf_final/final_model.rds")
 ```
@@
 ```{r}
+# Load in final model
+pokemon_tibble <- read_rds(file = "data/rf_final/final_model.rds")
+
 # Indicating the desired metric
 pokemon_metric <- metric_set(rmse)
-
-# Creating the predicted vs. actual value tibble
-pokemon_tibble <- predict(rf_final_fit, new_data = pokemon_test %>% select(-capture_rate))
-pokemon_tibble <- bind_cols(pokemon_tibble, pokemon_test %>% select(capture_rate))
 
@@
 pokemon_tibble_metrics
+
 ```
 
+Our random forest actually performed slightly better on the testing set than on the training set and folds with an RMSE of 42.42378!
+
 # Sources

### 2022-12-10 Final_Draft.Rmd
commit

@@
 
-Now, we will take that best model from the tuned random forest and fit it to the training data. This will train that random forest one more time on the entire training data set.
+Now, we will take that best model from the tuned random forest and fit it to the training data.
+This will train that random forest one more time on the entire training data set.
+Once we have fit and trained the random forest on the training data, it will be ready for testing!
 
@@
 # Save
-write_rds(rf_final_fit, file = "data/rf_final/rf_final_train.rds")
+write_rds(rf_final_fit_train, file = "data/rf_final/rf_final_train.rds")
 ```
-
-```{r}
-# Reading in the random forest fit on the training data
-rf_final_fit_train <- read_rds(file = "data/rf_final/rf_final_train.rds")
-
-# Collecting the RMSE of the random forest fit on the training data
-rf_final_fit_train %>% 
-  collect_metrics() %>% 
-  slice(1)
-```
-
-The chosen random forest model had an RMSE of 43.95341 on the training data. This is quite close to the RMSE the model had on the folds.
 
@@
 ```{r, eval = FALSE}
+# Loading in the training data fit
+rf_final_fit_train <- read_rds(file = "data/rf_final/rf_final_train.rds")
+
 # Creating the predicted vs. actual value tibble
@@
 pokemon_tibble_metrics
-
 ```
 
-Our random forest actually performed slightly better on the testing set than on the training set and folds with an RMSE of 42.42378!
+Our random forest actually performed better on the testing set than on the cross-validation folds with an RMSE of 37.87014!
+Since RMSE is measured in relation to the values the outcome may take on, our testing RMSE of 37.87014 in relation to the outcome range from 0 to 255 is not a great RMSE, but it is not too bad either.
+So, our model did not perform horribly and still managed to explain some variation in the outcome!
+
+## Variable Importance
+
+One cool thing about random forests is that they can also tell you which variables are the most important in predicting the outcome. This is shown in a variable importance plot (VIP).
+
+```{r}
+# Using the training fit to create the VIP because the model was not actually fit to the testing data
+rf_final_fit_train %>% 
+  extract_fit_engine() %>% 
+  vip(aesthetics = list(fill = "red3", color = "blue3"))
+```
+
+From this we can see that the variable that mattered the most in predicting the outcome was `base_total`, which is what we had believed would happen.
+Although some of the other variables present importance in predicting the outcome, the `base_total` is by far the most important variable.
+This also makes contextual sense because a Pokemon with higher battle statistics should be harder to capture because it has greater power than other Pokemon.
 

### 2022-12-11 Final_Draft.Rmd
commit

@@
 # RANDOM FOREST
-rf_parameter_grid <- grid_regular(mtry(range = c(1, 14)), trees(range = c(200,1000)), min_n(range = c(5,20)), levels = 8)
+rf_parameter_grid <- grid_regular(mtry(range = c(1, 12)), trees(range = c(200,1000)), min_n(range = c(5,20)), levels = 8)
 
@@
 For the random forest, we tuned the the minimal node size, the number of randomly selected predictors, and the number of trees.
-The number of trees does not seem to have much effect on the performance of the model.
+It may be important to note that the range of the number of randomly selected predictors for this model goes up to 12 rather than 14 (all predictors) to avoid creating a bagging model.
+The problem with a bagging model is that, because all the predictors are used, each tree may make the same first split.
+If all the trees have the same first split, every tree would no longer be independent from one another, which is an important assumption we make.
+And so, the range of the number of randomly selected predictors should be less than the total number of predictors, which is why I chose a slightly lower value at 12.
+From the plots, we can see that the number of trees does not seem to have much effect on the performance of the model.
 The minimal node size does not appear to have a huge effect on the performance, but the smaller values of `min_n` appear to have slightly lower RMSE values (at least judging by the starting point).
@@
 From the plots, it appears that, overall, a greater number of predictors renders a better performance.
-Actually, it seems that including all 14 predictors gives the lowest RMSE here.
+It seems that including 10-12 predictors gives the lowest RMSE here.
 Based on the RMSE values in the plots, this model definitely performs the best!
@@
 
-Random forest #40 with all 14 predictors, 657 trees, and a minimal node size of 5 performed the best with an RMSE of 43.79638!
+Random forest #31 with 10 predictors, 542 trees, and a minimal node size of 5 performed the best with an RMSE of 43.89281!
 
@@
 
-Our random forest actually performed better on the testing set than on the cross-validation folds with an RMSE of 37.87014!
-Since RMSE is measured in relation to the values the outcome may take on, our testing RMSE of 37.87014 in relation to the outcome range from 0 to 255 is not a great RMSE, but it is not too bad either.
+Our random forest actually performed better on the testing set than on the cross-validation folds with an RMSE of 37.78122!
+Since RMSE is measured in relation to the values the outcome may take on, our testing RMSE of 37.78122 in relation to the outcome range from 0 to 255 is not a great RMSE, but it is not too bad either.
 So, our model did not perform horribly and still managed to explain some variation in the outcome!
@@
 
+# Conclusion
+
+After fitting various models and conducting analysis on each of them, the best model to predict the capture rate of a Pokemon seems to be the random forest model.
+However, this model still does not do a great job predicting the capture rate (based on the RMSE).
+This might mean the capture rate of a Pokemon is not very predictable or that the predictors chosen might not have been the best predictors to use.
+
+Some avenues of improvement may be to look into other characteristics associated with capturing a Pokemon that may help explain more of the variability in the outcome than our current model.
+For example, in the Pokemon games, there are multiple Poke balls that can be used to capture a Pokemon (Great Ball, Ultra Ball, etc.).
+Perhaps the capture rate of some of the Pokemon in this data set was affected by which ball was used to capture it since some balls are more powerful than others.
+Another potential predictor could be the items used on a Pokemon before capturing it that might make it easier to capture (berries, etc.).
+This might have also altered the capture rate of some Pokemon.
+
+One potential question that may be lingering is whether the choice to include the `base_total` rather than the individual six battle statistics was the correct one.
+You may be wondering whether the model would have done better if those were included instead.
+I went back and included those six battle statistics instead of the `base_total` and ran the random forest model.
+The model actually did slightly worse than it did with the total of the statistics.
+When running the variable importance plot for that fit, `height_m` actually came up as the most important variable.
+I believe this might be due to the lower correlation each of the individual six statistics have with `capture_rate`.
+Perhaps any one of the individual six statistics alone does not contribute much to predicting the outcome.
+So, I believe it was justified to use `base_total` instead.
+
+If I were to continue this project and move forward with my analysis, I would like to explore more predictors that could contribute to finding the capture rate of a Pokemon (as mentioned before).
+Also, another avenue for analysis might be the Pokemon GO games instead of the original Pokemon games.
+It would be interesting to see if the capture rate of a Pokemon differs based on the game it is in or if the ability to predict the capture rate differs by game type.
+
+Overall, attempting to predict the capture rate of a Pokemon using this data set provided great opportunity for me to build my machine learning and data analysis skills.
+As I got to know the data set more, I found myself becoming more passionate about learning everything about my data and finding a model that would work the best.
+Although the random forest might not have been perfect, I am glad I was able to develop a model that at least explains some of the variation in the capture rate of a Pokemon!
+
 # Sources

### 2022-12-11 Final_Draft.Rmd
commit

@@
 
+Here is a visualization of these results:
+
+```{r}
+# Creating a data frame of the model RMSE's so we can plot
+all_models <- data.frame(Model = c("Linear Regression", "Ridge Regression", "Lasso Regression", "Polynomial Regression", "K Nearest Neighbors", "Elastic Net", "Random Forest", "Boosted Trees"),
+                         RMSE = c(lm_rmse$mean, ridge_rmse$mean, lasso_rmse$mean, poly_rmse$mean, knn_rmse$mean, elastic_rmse$mean, rf_rmse$mean, boosted_rmse$mean))
+
+# Creating a barplot of the RMSE values
+ggplot(all_models, aes(x=Model, y=RMSE)) +
+  geom_bar(stat = "identity", aes(fill = Model)) +
+  scale_fill_manual(values = c("blue1", "red1", "blue2", "red2", "blue3", "red3", "blue4", "red4")) +
+  theme(legend.position = "none") 
+```
+
 From the performance of the models on the cross-validation data, we can see that the random forest performed the best!

### 2022-12-11 Final_Draft.Rmd
commit

@@
 author: "Shivani Kharva"
-date: "2022-12-06"
+date: "2022-12-11"
 output:
@@
     toc: true
+    toc_float: true
 ---
@@
 
-This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.
+This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon Database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.
 
@@
 From this definition, it seems likely that it would be harder to capture a Pokemon that is legendary than one that is not.
+Let's see if this is true or not.
 
@@
 7.
-Select the best tuned parameters for the model (based on RMSE), adjust the model workflow to include that version of the model, and fit the model with the best tuned parameters to the folds.
-Make sure to save the final fit.
+Collect the metrics of the tuned models, arrange in ascending order of mean to see what the lowest RMSE for that tuned model is, and slice to choose only the lowest RMSE.
+Save the RMSE to a variable for comparison.
 
 ```{r, eval=FALSE, warning=FALSE, message=FALSE}
-# select_best() to select the best tuned parameters for the model (based on RMSE)
-# finalize_workflow() to adjust the model workflow to include that version of the model
-# fit_resamples() to fit the model with the best tuned parameter to the folds
+# collect_metrics() to collect the RMSE
+# slice() to save only RMSE
 
 # LINEAR REGRESSION 
-# No need to select the best fit as linear regression had no tuning -> we just fit the one linear regression model we fit
+# Fitting the linear regression to the folds first (since it had no tuning)
 lm_fit <- fit_resamples(lm_workflow, resamples = pokemon_folds)
-write_rds(lm_fit, file = "data/final_models/lm_final.rds")
+lm_rmse <- collect_metrics(lm_fit) %>% 
+  slice(1)
 
 # RIDGE REGRESSION
-best_penalty <- select_best(ridge_tune, metric = 'rmse')
-ridge_final <- finalize_workflow(ridge_workflow, best_penalty)
-ridge_final_fit <- fit_resamples(ridge_final, resamples = pokemon_folds)
-write_rds(ridge_final_fit, file = "data/final_models/ridge_final.rds")
+ridge_rmse <- collect_metrics(ridge_tuned) %>% 
+  arrange(mean) %>% 
+  slice(49)
 
 # LASSO REGRESSION
-best_lasso_penalty <- select_best(tune_res2, metric = "rmse")
-lasso_final <- finalize_workflow(lasso_workflow, best_lasso_penalty)
-lasso_final_fit <- fit_resamples(lasso_final, resamples = pokemon_folds)
-write_rds(lasso_final_fit, file = "data/final_models/lasso_final.rds")
+lasso_rmse <- collect_metrics(lasso_tuned) %>% 
+  arrange(mean) %>% 
+  slice(35)
 
 # POLYNOMIAL REGRESSION
-best_poly<- select_best(poly_tuned, metric = "rmse")
-poly_final <- finalize_workflow(poly_wf, best_poly)
-poly_final_fit <- fit_resamples(poly_final, resamples = pokemon_folds)
-write_rds(poly_final_fit, file = "data/final_models/poly_final.rds")
+poly_rmse <- collect_metrics(poly_tuned) %>% 
+  arrange(mean) %>% 
+  slice(6)
 
 # K NEAREST NEIGHBORS
-best_knn <- select_best(knn_tune, metric = "rmse")
-knn_final <- finalize_workflow(knn_workflow, best_knn)
-knn_final_fit <- fit_resamples(knn_final, resamples = pokemon_folds)
-write_rds(knn_final_fit, file = "data/final_models/knn_final.rds")
+knn_rmse <- collect_metrics(knn_tuned) %>% 
+  arrange(mean) %>% 
+  slice(6)
 
 # ELASTIC NET
-best_elastic <- select_best(elastic_tune, metric = "rmse")
-elastic_final <- finalize_workflow(elastic_workflow, best_elastic)
-elastic_final_fit <- fit_resamples(elastic_final, resamples = pokemon_folds)
-write_rds(elastic_final_fit, file = "data/final_models/elastic_final.rds")
+elastic_rmse <- collect_metrics(elastic_tuned) %>% 
+  arrange(mean) %>% 
+  slice(73)
 
 # RANDOM FOREST
-best_rf <- select_best(rf_tuned)
-rf_final_workflow <- finalize_workflow(rf_workflow, best_rf)
-rf_final_fit <- fit_resamples(rf_final_workflow, resamples = pokemon_folds)
-write_rds(rf_final_fit, file = "data/final_models/rf_final.rds")
+rf_rmse <- collect_metrics(rf_tuned) %>% 
+  arrange(mean) %>% 
+  slice(513)
 
 # BOOSTED TREES
-best_boosted <- select_best(boosted_tuned)
-boosted_final_workflow <- finalize_workflow(boosted_workflow, best_boosted)
-boosted_final_fit <- fit_resamples(boosted_final_workflow, resamples = pokemon_folds)
-write_rds(boosted_final_fit, file = "data/final_models/boosted_final.rds")
-```
-
-8. Load in each of the best models and collect the RMSE's of each of the best model fits for comparison with one another.
-
-```{r, warning=FALSE, message=FALSE}
-# collect_metrics() to collect the RMSE
-# slice() to only collect the RMSE and not R squared
-
-# LINEAR REGRESSION 
-lm_final_fit <- read_rds(file = "data/final_models/lm_final.rds")
-lm_rmse <- collect_metrics(lm_final_fit) %>% slice(1)
-
-# RIDGE REGRESSION
-ridge_final_fit <- read_rds(file = "data/final_models/ridge_final.rds")
-ridge_rmse <- collect_metrics(ridge_final_fit) %>% slice(1)
-
-# LASSO REGRESSION
-lasso_final_fit <- read_rds(file = "data/final_models/lasso_final.rds")
-lasso_rmse <- collect_metrics(lasso_final_fit) %>% slice(1)
-
-# POLYNOMIAL REGRESSION
-poly_final_fit <- read_rds(file = "data/final_models/poly_final.rds")
-poly_rmse <- collect_metrics(poly_final_fit) %>% slice(1)
-
-# K NEAREST NEIGHBORS
-knn_final_fit <- read_rds(file = "data/final_models/knn_final.rds")
-knn_rmse <- collect_metrics(knn_final_fit) %>% slice(1)
-
-# ELASTIC NET
-elastic_final_fit <- read_rds(file = "data/final_models/elastic_final.rds")
-elastic_rmse <- collect_metrics(elastic_final_fit) %>% slice(1)
-
-# RANDOM FOREST
-rf_final_fit <- read_rds(file = "data/final_models/rf_final.rds")
-rf_rmse <- collect_metrics(rf_final_fit) %>% slice(1)
-
-# BOOSTED TREES
-boosted_final_fit <- read_rds(file = "data/final_models/boosted_final.rds")
-boosted_rmse <- collect_metrics(boosted_final_fit) %>% slice(1)
+boosted_rmse <- collect_metrics(boosted_tuned) %>% 
+  arrange(mean) %>% 
+  slice(126)
 ```
@@
 ```
+*Note: the colors in this bar plot do not represent anything and are simply for aesthetic preference.*
 
@@
 # Fitting to the training data
-best_rf_train <- select_best(rf_tuned)
+best_rf_train <- select_best(rf_tuned, metric = 'rmse')
 rf_final_workflow_train <- finalize_workflow(rf_workflow, best_rf)
@@
 
+We might also be interested in a plot of the predicted values versus the actual values:
+
+```{r}
+# Creating plot of predicted values vs. actual values
+pokemon_tibble %>% 
+  ggplot(aes(x = .pred, y = capture_rate)) +
+  geom_point(alpha = 0.4) +
+  geom_abline(lty = 2) +
+  theme_grey() +
+  coord_obs_pred()
+```
+
+If every observation was predicted accurately, the dots would form a straight line.
+From the plot, we can see that very few of the dots actually fall on the line.
+However, it also appears that the model did not do a terrible job predicting the capture_rate because the points do seem to follow the same general pattern as the line.
+Also, it is a good sign that the model did not predict any negative capture rate values.
+So, overall, the random forest model did not do a terrible job predicting the capture rate, but there is definitely room for improvement.
+
 ## Variable Importance
@@
 After fitting various models and conducting analysis on each of them, the best model to predict the capture rate of a Pokemon seems to be the random forest model.
+This is no surprise at all since the random forest tends to work well for most data because it is nonparametric and makes no assumptions about parametric forms or the outcome (more flexible).
 However, this model still does not do a great job predicting the capture rate (based on the RMSE).
-This might mean the capture rate of a Pokemon is not very predictable or that the predictors chosen might not have been the best predictors to use.
+
+The model that did the worst was K Nearest Neighbors (KNN).
+This was also not surprising as KNN tends to do worse when there are too many predictors because this means the data has too many dimensions.
+In a high dimensional data space, the data points are not close enough to each other for KNN to do well in predicting the outcome unless there are enough observations to make up for the high dimensional data space.
+Therefore, it makes sense that KNN did the worst.
 
@@
 # Sources
+
+This data was taken from the Kaggle data set, "[Pokemon Stats Dataset](https://www.kaggle.com/datasets/hasanarcas/pokemon-stats-dataset)," and it was scraped from the [official Pokemon Database](https://pokemondb.net/pokedex/all) by user Hasan Arcas.
+
+The information about Minior was found on [Bulbapedia](https://bulbapedia.bulbagarden.net/wiki/Minior_(Pok%C3%A9mon)), a community Pokemon encyclopedia.
+
+General facts and definitions mentioned in the project were either found or double checked (if from prior knowledge) on the [official Pokemon Database](https://pokemondb.net/).
+This is also the same database where the data was scraped from.

### 2022-12-12 Predicting_Pokemon_Capture_Rate.Rmd
finishing up

--- 2022-12-11 Final_Draft.Rmd
@@
     toc_float: true
+    df_print: paged
+    theme: journal
 ---
@@
 The purpose of this project is to develop a model that will predict the capture rate of a Pokemon.
+
+<center>![](https://media.tenor.com/8vuqpD3Ir-IAAAAC/catching-pokemon.gif)</center>
 
@@
 
+<center>![Pokemon in battle!](https://static.wikia.nocookie.net/essentialsdocs/images/7/70/Battle.png/revision/latest?cb=20220523172438)</center>
+
 ## What Are We Trying to Do?
@@
 So, I would like to test if it is possible to develop a model that can predict the capture rate of a Pokemon based on those qualities.
+
+<center>![Pokemon stats screen.](https://78.media.tumblr.com/76016a9930e876e104b528190af0e125/tumblr_oxbo3x64Ge1w5miaho1_1280.png)</center>
 
@@
 library(tidytext)
+library(ggplot2)
+theme_set(theme_bw())
 
@@
 
+<center>![Left: Minior with its shell (Meteorite). Right: Minior without its shell (Core).](https://i1.wp.com/hallowpeakgaming.com/wp-content/uploads/2016/12/minior.jpg?fit=1024%2C486&ssl=1)</center>
+
 ```{r}
@@
 Now the capture rate for Minior is fixed and our outcome is an integer we can use in regression!
+
+<center>![](https://images-wixmp-ed30a86b8c4ca887773594c2.wixmp.com/f/2aaacf79-3286-4339-9843-f60dfc484cc3/dety9on-74a91eb7-7f6a-43e7-a87f-67bcb6a4796b.gif?token=eyJ0eXAiOiJKV1QiLCJhbGciOiJIUzI1NiJ9.eyJzdWIiOiJ1cm46YXBwOjdlMGQxODg5ODIyNjQzNzNhNWYwZDQxNWVhMGQyNmUwIiwiaXNzIjoidXJuOmFwcDo3ZTBkMTg4OTgyMjY0MzczYTVmMGQ0MTVlYTBkMjZlMCIsIm9iaiI6W1t7InBhdGgiOiJcL2ZcLzJhYWFjZjc5LTMyODYtNDMzOS05ODQzLWY2MGRmYzQ4NGNjM1wvZGV0eTlvbi03NGE5MWViNy03ZjZhLTQzZTctYTg3Zi02N2JjYjZhNDc5NmIuZ2lmIn1dXSwiYXVkIjpbInVybjpzZXJ2aWNlOmZpbGUuZG93bmxvYWQiXX0.rFMKqBaSDLqE8knqyrMLA-DOqL7KOLTl7iawmzojIyI)</center>
 
@@
 
+<center>![](https://media.tenor.com/tEGO5Q-EQIcAAAAC/pokemon.gif)</center>
+
 # Exploratory Data Analysis
@@
 
+Now, we will explore the relationships between select variables with the outcome as well as with each other!
+
 ### Capture Rate
@@
 ggplot(pokemon_data, aes(capture_rate)) +
-  geom_bar(fill='red3')
+  geom_bar(fill='red3') +
+  labs(
+    title = "Distribution of Capture Rate"
+  )
 ```
@@
 The possible values of `capture_rate` range from 0 to 255 and the `capture_rate` of the Pokemon in our data appears to range from 3 to 255.
-There is also a peak at about 45, meaning the most common capture rate for a Pokemon in our data is 45.
+There is also a peak at about 45, meaning the most common capture rate for a Pokemon in our data is ~45.
 
@@
 This poses the question of whether there are specific generations that have higher capture rates overall than any other generations.
+Also, legendary Pokemon are especially powerful Pokemon that players rarely encounter.
+From this definition, it seems likely that it would be harder to capture a Pokemon that is legendary than one that is not.
+Let's check to see how these variables relate to one another and to the capture rate!
 
 ```{r}
-# Boxplot of generation vs. capture rate
-ggplot(pokemon, aes(x=generation, y=capture_rate, group=generation)) +
-  geom_boxplot(aes(fill = generation))
+# Boxplot of generation and legendary status vs. capture rate
+pokemon_data %>% 
+  ggplot(aes(x = capture_rate, y = reorder(is_legendary, capture_rate), fill = generation)) +
+  geom_boxplot() +
+  labs(
+    title = "Box Plot of Generation and Legendary vs. Capture Rate",
+    y = "Legendary Status", 
+    x = "Capture Rate") +
+  scale_fill_manual(values = c("red1", "blue1", "red3", "blue3", "darkred", "blue4", "#660000"))
 ```
@@
 
-### Legendary Status
+It is also apparent that there are very few Pokemon, if any, of legendary status in any generations other than 6 and 7.
+However, the distribution of generations 6 and 7 across legendary Pokemon still fit the aforementioned observations.
 
-Legendary Pokemon are especially powerful Pokemon that players rarely encounter.
-From this definition, it seems likely that it would be harder to capture a Pokemon that is legendary than one that is not.
-Let's see if this is true or not.
+As for legendary status, from the plot, we can see that, although there are still non-legendary Pokemon in the lower end of the capture rate, there are no legendary Pokemon with a capture rate above 50.
+All legendary Pokemon capture rates appears to be less than all first quartiles of non-legendary Pokemon capture rates.
+The data implies that legendary Pokemon genearally have a lower capture rate than non-legendary Pokemon, which is what was expected.
+
+### Type 1 Plot
+
+In our analysis, we are only using Type 1 of each Pokemon since not all Pokemon have a Type 2. However, how many types of Pokemon are there?
 
 ```{r}
-ggplot(pokemon_data, aes(capture_rate)) + 
-  geom_bar(aes(fill = is_legendary)) +
-  scale_fill_manual(values = c('red2', 'blue4'))
+# Creating a bar chart of `type1`
+pokemon_data %>% 
+  ggplot(aes(y = type1)) +
+  geom_bar(fill = "#003399") +
+  labs(
+    title = "Count of Each Type"
+  )
 ```
 
-From the graph, we can see that, although there are still non-legendary Pokemon in the lower end of the capture rate, there are no legendary Pokemon with a capture rate above 50.
-Also, the lowest capture rate in our data, 3, is made up of almost all legendary Pokemon.
-It seems that legendary Pokemon have a lower capture rate than non-legendary Pokemon, which is what was expected.
+From the plot, we can see that there are 18 different types for type 1 Pokemon.
+18 types is a lot of categories to include!
+However, the plot also reveals that there are significantly more Pokemon in some types like "water" and "normal" in comparison to other types with much fewer Pokemon like "flying" and "fairy".
+So, in our analysis we should specify that we only want to prioritize the types with the most Pokemon and group the rest of the types into "Other".
+This will be completed and further discussed in the recipe creation.
 
@@
 # Visualization of correlation matrix
-pokemon_corrplot <- corrplot(pokemon_cor, method = "number")
+pokemon_corrplot <- corrplot(pokemon_cor, method = "circle", addCoef.col = 1, number.cex = 0.7)
 ```
 
-One relationship that stands out is the relatively strong negative correlation between `base_total` and `capture_rate` (-0.71), which indicates that it may be useful to use `base_total` in our analysis rather than the six individual battle statistics (we cannot use all of them because the battle statistics add up to the exact `base_total`).
+One relationship that stands out is the moderately negative correlation between `base_total` and `capture_rate` (-0.71), which indicates that it may be useful to use `base_total` in our analysis rather than the six individual battle statistics (we cannot use all of them because the battle statistics add up to the exact `base_total`).
 This relationship implies that, as a Pokemon has a higher total of battle statistics, it becomes more difficult to capture them (lower `capture_rate`).
@@
 Since the battle statistics have a similar relationship with `capture_rate` as `base_total`, this further justifies using only `base_total` rather than the six individual battle statistics.
+
+### Base Total
+
+Since we discussed the relationship between `base_total` and `capture_rate` in the correlation plot, perhaps we should further explore the relationship between the two variables.
+
+```{r, message = FALSE, warning=FALSE}
+pokemon_data %>% 
+  ggplot(aes(x=base_total, y=capture_rate)) + 
+  geom_jitter(width = 0.5, size = 1) +
+  geom_smooth(method = "lm", se =F, col="darkred") +
+  labs(title = "Capture Rate vs. Base Total")
+```
+
+Although the relationship is not very strong, it is apparent that there is a moderately negative relationship between `capture_rate` and `base_total`.
+That is, a higher value of total battle statistics slightly relates to a lower capture rate.
+This makes contextual sense because a Pokemon with greater battle abilities would be more valuable (and harder to capture because of its greater value).
+So, it seems as if it would prove more meaningful to use `base_total` rather than the six battle statistics in exploring the relationship between the predictors and the outcome, `capture_rate`.
 
@@
 
-We will use k-fold stratified cross validation in order to train the data on 10 mini datasets (or folds) created using the training data.
-Since we will be essentially using 10 mini training/testing sets to train our data, this will give us a better estimate of each model's true performance.
+We will create 10 folds to conduct k-fold (10-fold in our case) stratified cross validation.
+This means that R is taking the training data and assigning each observation in the training data to 1 of 10 folds.
+For each fold, a testing set is created consisting of that fold and the remaining k-1 folds will be the training set for that fold.
+At the end, we end up with k total folds.
+
+K-fold cross validation is done by splitting the data into k folds as described above with each fold being a testing set with the other k-1 folds being the training set for that fold.
+Then, whichever model we are fitting is fit to each training set and tested on the corresponding testing set (each time, a different fold should be used as a validation set).
+Then, the average accuracy is taken from the testing set of each of the folds to measure performance (other metrics can be taken as well such as standard error).
+
+We use k-fold cross validation rather than simply fitting and testing models on the entire training set because cross validation provides a better estimate of the testing accuracy.
+It is better to take the mean accuracy from several samples instead of just one accuracy from one sample because, as n increases, we reduce variation.
+
 We stratify on the outcome, `capture_rate`, to make sure the data in each fold is not imbalanced.
@@
 The RMSE is one of the most commonly used measures for evaluating the performance of regression models by showing how far the model's predictions are from the true values using Euclidian distance.
+So, a lower RMSE is better since that means the predicted values have a smaller distance from the actual values.
 Since RMSE measures distance, it is important that we normalize our data, which we did in the recipe.
@@
 Let's get to building our models!
+
+<center>![](https://pbs.twimg.com/media/EYoS4TmXgAIfLRD.jpg)</center>
 
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE}
+```{r, warning=FALSE, message=FALSE}
 # LINEAR REGRESSION 
@@
 lasso_workflow <- workflow() %>% 
-  add_recipe(ridge_lasso_recipe) %>% 
+  add_recipe(pokemon_recipe) %>% 
   add_model(lasso_spec)
@@
 
-```{r, eval=FALSE, warning=FALSE, message=FALSE}
+```{r, warning=FALSE, message=FALSE}
 # collect_metrics() to collect the RMSE
@@
   scale_fill_manual(values = c("blue1", "red1", "blue2", "red2", "blue3", "red3", "blue4", "red4")) +
-  theme(legend.position = "none") 
+  theme(legend.position = "none") +
+  labs(title = "Comparing RMSE by Model")
 ```
+
 *Note: the colors in this bar plot do not represent anything and are simply for aesthetic preference.*
@@
 One key thing to note from these results is that it appears the linear and simpler models did the worst, which indicates that our data is probably nonlinear.
+
+<center>![](https://media4.giphy.com/media/6nWhy3ulBL7GSCvKw6/giphy.gif)</center>
 
@@
 This means that the model does better when it is learning faster.
-Also, it appears that generally, once there are more than ~50 trees, the number of trees does not make much of a change on the performance of the model.
-However, the model appears to do worse when there are less than ~50 trees.
+Also, it appears that generally, once there are more than \~50 trees, the number of trees does not make much of a change on the performance of the model.
+However, the model appears to do worse when there are less than \~50 trees.
 The minimal node size does not seem to have much of an effect on the performance of the model.
@@
 Random forest #31 with 10 predictors, 542 trees, and a minimal node size of 5 performed the best with an RMSE of 43.89281!
+
+<center>![](https://www.ungeek.ph/wp-content/uploads/2016/11/ash_sparkiling-eyes.gif)</center>
 
@@
   theme_grey() +
-  coord_obs_pred()
+  coord_obs_pred() +
+  labs(title = "Predicted Values vs. Actual Values")
 ```
@@
 ```{r}
+# Loading in the training data fit
+rf_final_fit_train <- read_rds(file = "data/rf_final/rf_final_train.rds")
+
 # Using the training fit to create the VIP because the model was not actually fit to the testing data
@@
 
+<center>![](https://i.pinimg.com/originals/8f/23/e4/8f23e48ae48e3a07b342a718249e0080.gif)</center>
+
 # Sources

### 2023-04-06 Predicting_Pokemon_Capture_Rate.Rmd
final commit
edits

@@
 title: "Predicting the Capture Rate of Pokemon"
-subtitle: "PSTAT 131 Final Project"
 author: "Shivani Kharva"
@@
     code_folding: hide
-    toc: true
-    toc_float: true
+    toc: yes
+    toc_float: yes
     df_print: paged
     theme: journal
+  pdf_document:
+    toc: yes
+subtitle: PSTAT 131 Final Project
 ---
@@
 
-However, a question I'd like to pose is "How hard is to capture a Pokemon in the first place depending on their existing statistics?" It seems quite reasonable to hypothesize that a Pokemon's capture rate would vary depending on variables like their type, base statistics, generation, etc.
+However, a question I'd like to pose is "How hard is it to capture a Pokemon in the first place depending on their existing statistics?" It seems quite reasonable to hypothesize that a Pokemon's capture rate would vary depending on variables like their type, base statistics, generation, etc.
 So, I would like to test if it is possible to develop a model that can predict the capture rate of a Pokemon based on those qualities.
@@
 
-One last thing we need to do before our data is tidy the `abilities` variable because we currently have a column made up of lists of each Pokemon's abilities.
+One last thing we need to do before our exploratory data analysis is tidy the `abilities` variable because we currently have a column made up of lists of each Pokemon's abilities.
 However, it is not very useful to analyze 801 unique lists of abilities.
@@
 
-### Type 1 Plot
+### Type 1
 
